---
title: "This is where you can find some VAERS analysis - lots of pretty pictures..."
author: "Dr. Jessica Rose"
date: "2021-05-21"
output: html_document
---

This is an R Markdown sheet generated from the VAERS data. For the most part, only figures are shown. I decided to show figures pertaining to Death, Female Reproductive Issues, Breakthrough COVID-19 infections and Cardiovascular and Immunological adverse events. ALso included is some stuff on kids and Myocarditis.

I start by coalescing the .csv files downloaded from the VAERS website: https://vaers.hhs.gov/data/datasets.html

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=5, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```
```{r}
library("readr")
library("tidyverse")
library("dplyr")
library("ggplot2")
library("stringr")
library("DT")
library("reshape2")
library("data.table")
library("plotly")
library("xtable")
library("blogdown")
```
This is the total number of adverse events in the context of COVID-19 products from Fizer, Modrna and JJ
```{r}
VAERSSYM_01_10_21_490500kb <- read_csv("/Users/jessicarose/Desktop/2021VAERSData_01_10_21_490500kb/2021VAERSSYMPTOMS.csv", col_types = cols(VAERS_ID = col_double()))
VAERSDATA_01_10_21_490500kb <- read_csv("/Users/jessicarose/Desktop/2021VAERSData_01_10_21_490500kb/2021VAERSDATA.csv", col_types = cols(VAERS_ID = col_double()))
VAERSVAX_01_10_21_490500kb <- read_csv("/Users/jessicarose/Desktop/2021VAERSData_01_10_21_490500kb/2021VAERSVAX.csv", col_types = cols(VAERS_ID = col_double()))
RV_VAX_01_10_21_490500kb <- VAERSVAX_01_10_21_490500kb %>% filter(VAX_TYPE == "COVID19")
MERGED_SYM_DAT_01_10_21_490500kb <- full_join(VAERSSYM_01_10_21_490500kb, VAERSDATA_01_10_21_490500kb, by = c("VAERS_ID"))
MERGED_SYM_DAT_VAX_01_10_21_490500kb <- full_join(MERGED_SYM_DAT_01_10_21_490500kb, RV_VAX_01_10_21_490500kb, by = c("VAERS_ID"))
MERGED_01_10_21_490500kb_1 <- MERGED_SYM_DAT_VAX_01_10_21_490500kb %>% distinct(VAERS_ID, .keep_all = TRUE)
MERGED_01_10_21_490500kb <- MERGED_01_10_21_490500kb_1 %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)
```
```{r}
#write.csv(MERGED_01_10_21_490500kb, "/Users/jessicarose/Desktop/MERGED_01_10_21_490500kb.csv")
```
```{r}
MERGED_01_10_21_490500kb_Y <- read_csv("/Users/jessicarose/Desktop/MERGED_01_10_21_490500kb_Y.csv")
```
```{r}
MERGED_01_10_21_490500kb <- MERGED_01_10_21_490500kb_Y
```
```{r}
MERGED_01_10_21_490500kb$VAX_DATE <- as.Date(MERGED_01_10_21_490500kb$VAX_DATE, "%m/%d/%y") 
MERGED_01_10_21_490500kb$ONSET_DATE <- as.Date(MERGED_01_10_21_490500kb$ONSET_DATE, "%m/%d/%y")
MERGED_01_10_21_490500kb$DATEDIED <- as.Date(MERGED_01_10_21_490500kb$DATEDIED, "%m/%d/%y")
MERGED_01_10_21_490500kb$TODAYS_DATE <- as.Date(MERGED_01_10_21_490500kb$TODAYS_DATE, "%m/%d/%y")
MERGED_01_10_21_490500kb$RECVDATE <- as.Date(MERGED_01_10_21_490500kb$RECVDATE, "%m/%d/%y")
nrow(MERGED_01_10_21_490500kb)
```
Percent of VAERS reports made by Females
```{r}
MERGED_01_10_21_490500kb_SEX_no_U <- MERGED_01_10_21_490500kb %>% filter(SEX != "U")
MERGED_01_10_21_490500kb_SEX_no_U_F <- MERGED_01_10_21_490500kb_SEX_no_U %>% filter(SEX == "F")
nrow(MERGED_01_10_21_490500kb_SEX_no_U_F)/nrow(MERGED_01_10_21_490500kb)*100
```
The following table is a summary of some variables in the context of COVID-19 product roll-out including changing fully-injected numbers (FV = Fully Vaccinated; SAE = Severe Adverse Events; ER = Emergency doctor visits; COVID = Breakthrough infections; FRI = Female Reproductive Issues)
```{r}
Table <- data.frame(
  DATE = as.Date(c("1/16/21", "1/23/21", "1/30/21", "2/13/21", "2/27/21", "3/5/21", "3/12/21", "3/19/21", "3/26/21", "4/2/21", "4/9/21", "4/16/21", "4/23/21", "4/30/21", "5/7/21", "5/14/21", "5/21/21", "5/28/21", "6/4/21", "6/11/21", "6/18/21", "6/25/21", "7/2/21", "7/9/21", "7/16/21", "7/23/21", "7/30/21", "8/6/21", "8/13/21", "8/20/21", "8/27/21", "9/3/21", "9/10/21", "9/17/21", "9/24/21", "10/1/21"), "%m/%d/%y"),
  IDs = as.numeric(c("1431", "2160", "2946", "5351", "9286", "14701", "20586", "27955", "34121", "40348", "46163", "57641", "75370", "108191", "146622", "182559", "217716", "252523", "284041", "316929", "346967", "368162", "388900", "397262", "403335", "409924", "419295", "427831", "436175", "450602", "473722", "499779", "521667", "541594", "551836", "565282")),
  FV = as.numeric(c("4582089", "7664179", "11037313", "18895522", "27167910", "31720149", "35000000", "44145522", "48700000", "57980000", "64420000", "80610000", "89250000", "99670000", "108930000", "118990000", "126610000", "132770000", "136640000", "141580000", "148460000", "151620000", "155880000", "158290000", "160410000", "162170000", "163870000", "165640000", "167350000", "169590000", "172170000", "174120000", "177430000", "180570000", "182590000", "184340000")),
  DEATH = as.numeric(c("137", "281", "456", "810", "984", "1162", "1419", "1561", "1957", "2149", "2240", "2500", "3084", "3442", "3731", "4015", "4169", "4372", "4547", "4700", "4825", "4931", "5163", "5375", "5508", "5604", "6538", "6639", "6841", "7018", "7149", "7344", "7662", "7859", "8091", "8342")),
  HOSPITAL = as.numeric(c("338", "607", "953", "1747", "2195", "2676", "3412", "3913", "4387", "4758", "4906", "5009", "6215", "8099", "10648", "11557", "12625", "14977", "17520", "19354", "20571", "21636", "22612", "23441", "24087", "24778", "25634", "26402", "27160", "27913", "28651", "29577", "31041", "32024", "33172", "34288")), 
  ER = as.numeric(c("338", "576", "847", "1493", "2187", "2980", "3855", "4763", "5529", "6329", "6983", "8084", "10375", "14553", "19620", "23641", "27774", "32547", "37184", "41937", "45952", "49511", "52399", "53789", "55016", "56357", "57820", "59061", "60317", "61754", "63115", "64893", "66657", "68692", "70166", "71831")),
  SAE = as.numeric(c("690", "1207", "1806", "3221", "4291", "5535", "7065", "8470", "9796", "10930", "11724", "13057", "16478", "22145", "29024", "33791", "38650", "44860", "51068", "57073", "61917", "66255", "69941", "71963", "73645", "75536", "78186", "80011", "81937", "84137", "86209", "88768", "91839", "94740", "97147", "99838")),
  COVID = as.numeric(c("194", "254", "326", "619", "717", "776", "943", "1057", "1169", "1267", "1375", "1404", "1587", "1930", "2776", "3317", "3742", "4285", "4850", "5497", "6151", "6927", "8324", "9139", "9671", "10199", "10728", "11381", "12174", "13117", "14448", "15979", "17626", "18917", "19953", "21042")), 
  FRI = as.numeric(c("4", "7", "8", "24", "34", "43", "67", "77", "88", "98", "105", "123", "165", "287", "436", "597", "774", "1067", "1421", "1967", "2460", "2940", "4145", "4934", "5723", "6130", "6538", "6639", "6841", "6904", "7105", "7405", "7722", "8303", "8646", "9153")),
  
  stringsAsFactors = FALSE
)
Table
```
```{r}
Table_IDs_percent <- data.frame(
  DATE = as.Date(c("1/16/21", "1/23/21", "1/30/21", "2/13/21", "2/27/21", "3/5/21", "3/12/21", "3/19/21", "3/26/21", "4/2/21", "4/9/21", "4/16/21", "4/23/21", "4/30/21", "5/7/21", "5/14/21", "5/21/21", "5/28/21", "6/4/21", "6/11/21", "6/18/21", "6/25/21", "7/2/21", "7/9/21", "7/16/21", "7/23/21", "7/30/21", "8/6/21", "8/13/21", "8/20/21", "8/27/21", "9/3/21", "9/10/21", "9/17/21", "9/24/21", "10/1/21"), "%m/%d/%y"),
  IDs = as.numeric(c("1431", "2160", "2946", "5351", "9286", "14701", "20586", "27955", "34121", "40348", "46163", "57641", "75370", "108191", "146622", "182559", "217716", "252523", "284041", "316929", "346967", "368162", "388900", "397262", "403335", "409924", "419295", "427831", "436175", "450602", "473722", "499779", "521667", "541594", "551836", "565282")),
  FV = as.numeric(c("4582089", "7664179", "11037313", "18895522", "27167910", "31720149", "35000000", "44145522", "48700000", "57980000", "64420000", "80610000", "89250000", "99670000", "108930000", "118990000", "126610000", "132770000", "136640000", "141580000", "148460000", "151620000", "155880000", "158290000", "160410000", "162170000", "163870000", "165640000", "167350000", "169590000", "172170000", "174120000", "177430000", "180570000", "182590000", "184340000")),
  DEATH = as.numeric(c("137", "281", "456", "810", "984", "1162", "1419", "1561", "1957", "2149", "2240", "2500", "3084", "3442", "3731", "4015", "4169", "4372", "4547", "4700", "4825", "4931", "5163", "5375", "5508", "5604", "6538", "6639", "6841", "7018", "7149", "7344", "7662", "7859", "8091", "8342")),
  HOSPITAL = as.numeric(c("338", "607", "953", "1747", "2195", "2676", "3412", "3913", "4387", "4758", "4906", "5009", "6215", "8099", "10648", "11557", "12625", "14977", "17520", "19354", "20571", "21636", "22612", "23441", "24087", "24778", "25634", "26402", "27160", "27913", "28651", "29577", "31041", "32024", "33172", "34288")), 
  ER = as.numeric(c("338", "576", "847", "1493", "2187", "2980", "3855", "4763", "5529", "6329", "6983", "8084", "10375", "14553", "19620", "23641", "27774", "32547", "37184", "41937", "45952", "49511", "52399", "53789", "55016", "56357", "57820", "59061", "60317", "61754", "63115", "64893", "66657", "68692", "70166", "71831")),
  SAE = as.numeric(c("690", "1207", "1806", "3221", "4291", "5535", "7065", "8470", "9796", "10930", "11724", "13057", "16478", "22145", "29024", "33791", "38650", "44860", "51068", "57073", "61917", "66255", "69941", "71963", "73645", "75536", "78186", "80011", "81937", "84137", "86209", "88768", "91839", "94740", "97147", "99838")),
  COVID = as.numeric(c("194", "254", "326", "619", "717", "776", "943", "1057", "1169", "1267", "1375", "1404", "1587", "1930", "2776", "3317", "3742", "4285", "4850", "5497", "6151", "6927", "8324", "9139", "9671", "10199", "10728", "11381", "12174", "13117", "14448", "15979", "17626", "18917", "19953", "21042")), 
  FRI = as.numeric(c("4", "7", "8", "24", "34", "43", "67", "77", "88", "98", "105", "123", "165", "287", "436", "597", "774", "1067", "1421", "1967", "2460", "2940", "4145", "4934", "5723", "6130", "6538", "6639", "6841", "6904", "7105", "7405", "7722", "8303", "8646", "9153")),
  
  stringsAsFactors = FALSE
)
#Table_IDs_percent
```

```{r}
SAE_calculations <- Table_IDs_percent %>% mutate("SAE_p_FV" = SAE/(FV/1000000)) %>% mutate("SAE_p_mill_doses" = SAE_p_FV/(2.1)) #cum num of doses/FV population = 331650000/158290000 = 2.1
#SAE_calculations
```

```{r}
IDs = as.numeric(c("1431", "2160", "2946", "5351", "9286", "14701", "20586", "27955", "34121", "40348", "46163", "57641", "75370", "108191", "146622", "182559", "217716", "252523", "284041", "316929", "346967", "368162", "388900", "397262", "403335", "409024", "419295", "427831", "436175", "450602", "473721", "499779", "521667", "541594", "551836", "565282"))
#(diff(IDs)/(IDs))*100
```

```{r}
Table_IDs_percent_FV_IDs <- Table_IDs_percent %>% mutate("DEATH_PERCENT_ID" = DEATH/IDs*100) %>% mutate("H_PERCENT_ID" = HOSPITAL/IDs*100) %>% mutate("ER_PERCENT_ID" = ER/IDs*100) %>% mutate("SAE_PERCENT_ID" = SAE/IDs*100) %>% mutate("COVID_PERCENT_ID" = COVID/IDs*100) %>%  mutate("DEATH_PERCENT_FV" = DEATH/FV*100) %>% mutate("H_PERCENT_FV" = HOSPITAL/FV*100) %>% mutate("ER_PERCENT_FV" = ER/FV*100) %>% mutate("SAE_PERCENT_FV" = SAE/FV*100) %>% mutate("COVID_PERCENT_FV" = COVID/FV*100) %>% mutate("FRI_PERCENT_FV" = FRI/FV*100)
#Table_IDs_percent_FV_IDs
```

```{r} 
IDs_normalized_total_IDs <- Table_IDs_percent %>% mutate("IDs_normal_ID" = IDs/nrow(MERGED_01_10_21_490500kb)*100)
#IDs_normalized_total_IDs
```

```{r}
#Table_IDs_percent_normalized_US_pop <- Table_IDs_percent_FV_IDs %>% mutate("IDs_norm_US_pop" = IDs/(US_population/1000000)) 
#Table_IDs_percent_normalized_US_pop

Table_IDs_percent_normalized_FV <- Table_IDs_percent_FV_IDs %>% mutate("IDs_norm_FV" = IDs/(FV/1000000)) 
#Table_IDs_percent_normalized_FV

Table_AEs_percent_normalized_per_million <- Table_IDs_percent_FV_IDs %>% mutate("Death_norm_per_million" = DEATH/(FV/1000000)) %>% mutate("Hospital_norm_per_million" = HOSPITAL/(FV/1000000)) %>% mutate("ER_norm_per_million" = ER/(FV/1000000)) %>% mutate("SAE_norm_per_million" = SAE/(FV/1000000)) %>% mutate("COVID_norm_per_million" = COVID/(FV/1000000)) %>% mutate("FRI_norm_per_million" = FRI/(FV/1000000))
#Table_AEs_percent_normalized_per_million
```

```{r}
Table_IDs_all_VAERS <- data.frame(
  DATE = as.double(c("1990", "1991", "1992", "1993", "1994", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021")),
  IDs = as.numeric(c("2787", "16657", "18392", "17655", "18441", "17794", "17592", "17247", "15997", "20072", "23113", "25492", "26852", "31248", "28663", "28664", "31353", "48349", "51019", "50524", "49601", "41200", "41149", "43335", "47165", "57224", "60450", "52751", "62305", "61204", "55833", "62000")),
  US_Pop = as.numeric(c("252120309", "254539370", "256990613", "259532129", "262241196", "265163745", "268335003" ,"271713635", "275175301", "278548150", "281710909", "284607993", "287279318", "289815562", "292354658", "294993511", "297758969", "300608429", "303486012", "306307567", "309011475", "311584047", "314043885", "316400538", "318673411", "320878310" ,"323015995", "325084756", "327096265", "329064917", "331002651", "332915073")),
  
  stringsAsFactors = FALSE
)
Table_IDs_all_VAERS <- Table_IDs_all_VAERS %>% mutate("NORMALIZED" = IDs/(US_Pop/1000000))
#head(Table_IDs_all_VAERS)
```
This is the change in absolute number of VAERS IDs per year.
```{r}
Table_IDs_all_VAERS_plot <- ggplot(data = Table_IDs_all_VAERS, aes(DATE, IDs, colour = "IDs")) + geom_line(data = Table_IDs_all_VAERS, aes(DATE, IDs, colour = "IDs"), colour = "red", size = 2) + geom_point(data = Table_IDs_all_VAERS, aes(DATE, IDs, colour = "IDs"), colour = "red", size = 5) + xlab("DATE") + ylab("ABSOLUTE NUMBER OF VAERS IDs (N)") + labs(title = "CHANGE IN ABSOLUTE VAERS ID COUNTS PER YEAR", caption = "Data source: VAERS", size = 12) + theme(axis.text.y = element_text(colour = "#000000", size = 14)) + theme(axis.text.x = element_text(angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + theme(legend.background = element_rect(fill = "white"), legend.key = element_rect(fill = "black", color = NA), legend.key.size = unit(0.5, "cm"), legend.key.width = unit(0.5,"cm")) + ylim(0,400000) #+ geom_smooth(method = "lm", se=FALSE, color="blue", aes(group=1))
Table_IDs_all_VAERS_plot
#ggsave("Table_IDs_all_VAERS_plot.png")
```
These are the VAERS ID counts for the past 30 years (absolute counts).
```{r}
#Table_IDs_all_VAERS_plot_norm_US <- ggplot(data = Table_IDs_all_VAERS, aes(DATE, NORMALIZED, colour = "NORMALIZED")) + geom_line(data = Table_IDs_all_VAERS, aes(DATE, NORMALIZED, colour = "NORMALIZED"), colour = "red", size = 2) + geom_point(data = Table_IDs_all_VAERS, aes(DATE, NORMALIZED, colour = "NORMALIZED"), colour = "red", size = 5) + xlab("DATE") +  ylab("RATE NORMALIZED TO U.S. POPLN/MILLION") + labs(title = "VAERS ID COUNTS FOR THE PAST 30 YEARS", caption = "Data source: VAERS", size = 12) + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + theme(legend.background = element_rect(fill = "white"), legend.key = element_rect(fill = "black", color = NA), legend.key.size = unit(0.5, "cm"), legend.key.width = unit(0.5,"cm")) + geom_smooth(method = "lm", se=FALSE, color="blue", aes(group=1)) #+ geom_text(aes(label = IDs), color = "white", size = 4) 
#Table_IDs_all_VAERS_plot_norm_US
#ggsave("Table_IDs_all_VAERS_plot_norm_US.png")
```
This is the cumulative number of VAERS IDs (N) so far for 2021 in the context of the COVID-19 products.
```{r}
Table_IDs_plot <- ggplot(data = Table_IDs_percent, aes(x = DATE, y = IDs)) + geom_line(data = Table_IDs_percent, aes(x = DATE, y = IDs, colour = "IDs"), colour = "red", size = 2) + geom_point(data = Table_IDs_percent, aes(DATE, IDs, colour = "IDs"), colour = "red", size = 5) + xlab("DATE") +  ylab("NUMBER OF VAERS IDs (N)") + labs(title = "CUMULATIVE CHANGES IN VAERS IDs (N)", caption = "Data source: VAERS", size = 12) + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=15,vjust=0.5,hjust=0.5, size = 8)) + theme(panel.background = element_rect(fill = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + scale_x_continuous(breaks = round(seq(min(Table_IDs_percent$DATE), max(Table_IDs_percent$DATE), by = 10),0.1)) #+ ylim(0,200000)
Table_IDs_plot
ggsave("Table_IDs_plot.png")
```

This is the number of VAERS IDs (N) so far for 2021 in the context of the COVID-19 products normalized to FULLY INJECTED population per million.
```{r}
Table_IDs_plot_FV_p_mill <- ggplot(data = Table_IDs_percent_normalized_FV, aes(x = DATE, y = IDs_norm_FV)) + geom_line(data = Table_IDs_percent_normalized_FV, aes(x = DATE, y = IDs_norm_FV, colour = "IDs_norm_FV"), colour = "red", size = 2) + geom_point(data = Table_IDs_percent_normalized_FV, aes(DATE, IDs_norm_FV, colour = "IDs_norm_FV"), colour = "red", size = 5) + xlab("DATE") +  ylab("RATE NORMALIZED TO U.S. FV POPLN/MILLION") + labs(title = "CHANGING RATES NORMALIZED TO U.S. FV POPLN/MILLION", caption = "Data source: VAERS", size = 12) + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=15,vjust=0.5,hjust=0.5, size = 8)) + theme(panel.background = element_rect(fill = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + scale_x_continuous(breaks = round(seq(min(Table_IDs_percent$DATE), max(Table_IDs_percent$DATE), by = 10),0.1)) #+ ylim(0,200000)
Table_IDs_plot_FV_p_mill
ggsave("Table_IDs_plot_FV_p_mill.png")
```
These are the trajectories of cumulative data for the respective groups. 
```{r}
Table_IDs_percent_plot <- ggplot() + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, DEATH, colour = "DEATH"), size = 2) + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, HOSPITAL, colour = "HOSPITAL"), size = 2) + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, ER, colour = "ER"), size = 2) + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, SAE, colour = "SAE"), size = 2) + xlab("DATE") +  ylab("NUMBER OF VAERS IDs (N)") + labs(title = "CUMULATIVE CHANGES (N) FOR SELECTED VARIABLES", caption = "Data source: VAERS", size = 12) + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, COVID, colour = "COVID"), size = 2) + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, FRI, colour = "FRI"), size = 2) + labs(caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + theme(legend.background = element_rect(fill = "white"), legend.key = element_rect(fill = "black", color = NA), legend.key.size = unit(0.5, "cm"), legend.key.width = unit(0.5,"cm")) + ylim(0,100000)
Table_IDs_percent_plot
ggsave("Table_IDs_percent_plot.png")
```
These are the trajectories of normalized data for the respective groups (normalized to US FULLY INJECTED population per million). 
```{r}
Table_IDs_FV_per_million <- ggplot() + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, Death_norm_per_million, colour = "DEATH"), size = 2) + xlab("DATE") +  ylab("RATE NORMALIZED TO U.S. FV POPLN/MILLION") + labs(title = "CHANGING RATES NORMALIZED TO U.S. FV POPLN/MILLION", caption = "Data source: VAERS", size = 12) + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, Hospital_norm_per_million, colour = "HOSPITAL"), size = 2) + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, ER_norm_per_million, colour = "ER"), size = 2) + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, SAE_norm_per_million, colour = "SAE"), size = 2) + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, COVID_norm_per_million, colour = "COVID"), size = 2) + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, FRI_norm_per_million, colour = "FRI"), size = 2) + labs(caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + theme(legend.background = element_rect(fill = "white"), legend.key = element_rect(fill = "black", color = NA), legend.key.size = unit(0.5, "cm"), legend.key.width = unit(0.5,"cm"))
Table_IDs_FV_per_million
ggsave("Table_IDs_FV_per_million.png")
```
```{r}
AGE_GROUPED_BY_DECADE <- function(AGE_YRS) {
  if(AGE_YRS >= 0 & AGE_YRS <10) {
    AGE_GROUP = "A. 0-10"
    } else if(AGE_YRS >= 10 & AGE_YRS <20) {
      AGE_GROUP = "B. 10-20"
      } else if(AGE_YRS >= 20 & AGE_YRS <30) {
        AGE_GROUP = "C. 20-30"
        } else if(AGE_YRS >= 30 & AGE_YRS <40) {
          AGE_GROUP = "D. 30-40"
          } else if(AGE_YRS >= 40 & AGE_YRS <50) {
            AGE_GROUP = "E. 40-50"
            } else if(AGE_YRS >= 50 & AGE_YRS <60) {
              AGE_GROUP = "F. 50-60"
              } else if(AGE_YRS >= 60 & AGE_YRS <70) {
                AGE_GROUP = "G. 60-70"
                } else if(AGE_YRS >= 70 & AGE_YRS <80) {
                  AGE_GROUP = "H. 70-80"
                  } else if(AGE_YRS >= 80 & AGE_YRS <90) {
                    AGE_GROUP = "I. 80-90"
                    } else if(AGE_YRS >= 90 & AGE_YRS <100) {
                      AGE_GROUP = "J. 90-100"
                      } else if(AGE_YRS >= 100 & AGE_YRS <110) {
                        AGE_GROUP = "K. 100-110"
                        }
  else {
    AGE_GROUP = "NA"
    }
  return(AGE_GROUP)
}
```
```{r}
MERGED_01_10_21_490500kb_AG <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
MERGED_01_10_21_490500kb_AG_count <- MERGED_01_10_21_490500kb_AG %>% count(AGE_GROUP)
MERGED_01_10_21_490500kb_count_plot <- ggplot(MERGED_01_10_21_490500kb_AG_count, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("ADVERSE EVENTS (N)") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("ALL AEs -> Distribution of AGE GROUPS")
MERGED_01_10_21_490500kb_count_plot
ggsave("MERGED_01_10_21_490500kb_count_plot.png")

MERGED_01_10_21_490500kb_AGE_count <- MERGED_01_10_21_490500kb %>% group_by(AGE_YRS) %>% filter(AGE_YRS != "NA") %>% count(AGE_YRS)
#MERGED_01_10_21_490500kb_AGE_count <- data.frame(MERGED_01_10_21_490500kb_AGE_count, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#MERGED_01_10_21_490500kb_AGE_count
```
This is the number of birth defects.
```{r}
MERGED_01_10_21_490500kb_doses <- MERGED_01_10_21_490500kb %>% group_by(VAX_DOSE_SERIES) %>% filter(VAX_DOSE_SERIES >= 0 & VAX_DOSE_SERIES <= 2) %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n))
#MERGED_01_10_21_490500kb_doses
MERGED_01_10_21_490500kb_doses_VM <- MERGED_01_10_21_490500kb %>% group_by(VAX_DOSE_SERIES, VAX_MANU) %>% filter(VAX_DOSE_SERIES >= 0 & VAX_DOSE_SERIES <= 2 & VAX_MANU != "UNKNOWN MANUFACTURER") %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n)) 
MERGED_01_10_21_490500kb_doses_n <- MERGED_01_10_21_490500kb_doses %>% mutate("SUM" = sum(MERGED_01_10_21_490500kb_doses$n)) %>% mutate("PERCENT" = n/SUM*100)
#MERGED_01_10_21_490500kb_doses_n
```
```{r}
BD_01_10_21_490500kb <- MERGED_01_10_21_490500kb %>% filter(BIRTH_DEFECT == TRUE)
nrow(BD_01_10_21_490500kb)
```
This is the number of deaths and the percentage per total AEs.
```{r}
D_01_10_21_490500kb <- MERGED_01_10_21_490500kb %>% group_by(VAERS_ID, DIED) %>% filter(DIED == "Y")
#nrow(D_01_10_21_490500kb)
#nrow(D_01_10_21_490500kb)/nrow(MERGED_01_10_21_490500kb)*100
```
```{r}
DEAD_MERGED_01_10_21_490500kb <- MERGED_01_10_21_490500kb %>% filter(DIED == "Y")
nrow(DEAD_MERGED_01_10_21_490500kb)
nrow(DEAD_MERGED_01_10_21_490500kb)/nrow(MERGED_01_10_21_490500kb)*100
```
These are the number of hospitalizations and emergency doctor visits. 
```{r}
HOSPITAL_01_10_21_490500kb  <- MERGED_01_10_21_490500kb  %>% group_by(VAERS_ID, HOSPITAL) %>% filter(HOSPITAL == "Y")
nrow(HOSPITAL_01_10_21_490500kb)

ER_ED_01_10_21_490500kb <- MERGED_01_10_21_490500kb %>% group_by(VAERS_ID, ER_ED_VISIT) %>% filter(ER_ED_VISIT == "Y")
nrow(ER_ED_01_10_21_490500kb)
```

These are the numbers of Severe Adverse Events (SAEs): Deaths, Hospital, ER, Life threatening, Disabling, Birth defect with the total SAE count and percentage of all AEs. 
```{r}
#MERGED_01_10_21_490500kb_Y_dead
HOSPITAL_Y_01_10_21_490500kb <- MERGED_01_10_21_490500kb %>% filter(HOSPITAL == "Y")
ER_ED_Y_01_10_21_490500kb <- MERGED_01_10_21_490500kb %>% filter(ER_ED_VISIT == "Y")
L_THREAT_Y_01_10_21_490500kb <- MERGED_01_10_21_490500kb %>% filter(L_THREAT == "Y")
DISABLE_Y_01_10_21_490500kb <- MERGED_01_10_21_490500kb %>% filter(DISABLE == "Y")
BIRTH_DEFECT_Y_01_10_21_490500kb <- MERGED_01_10_21_490500kb %>% filter(BIRTH_DEFECT == TRUE)

nrow(DEAD_MERGED_01_10_21_490500kb)
nrow(HOSPITAL_Y_01_10_21_490500kb)
nrow(ER_ED_Y_01_10_21_490500kb)
nrow(L_THREAT_Y_01_10_21_490500kb)
nrow(DISABLE_Y_01_10_21_490500kb)
nrow(BIRTH_DEFECT_Y_01_10_21_490500kb)

SAES_01_10_21_490500kb <- rbind(DEAD_MERGED_01_10_21_490500kb, HOSPITAL_Y_01_10_21_490500kb, ER_ED_Y_01_10_21_490500kb, L_THREAT_Y_01_10_21_490500kb, DISABLE_Y_01_10_21_490500kb, BIRTH_DEFECT_Y_01_10_21_490500kb)

SAES_no_dups_01_10_21_490500kb <- subset(SAES_01_10_21_490500kb, !duplicated(subset(SAES_01_10_21_490500kb, select=c(VAERS_ID))))
nrow(SAES_no_dups_01_10_21_490500kb)
Percent = nrow(SAES_no_dups_01_10_21_490500kb)/nrow(MERGED_01_10_21_490500kb)*100
Percent
```

Histogram showing distribution of SAEs according to age group
```{r}
SAES_no_dups_01_10_21_490500kb_AG <- SAES_no_dups_01_10_21_490500kb %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
SAES_no_dups_01_10_21_490500kb_AG_count <- SAES_no_dups_01_10_21_490500kb_AG %>% count(AGE_GROUP)
SAES_no_dups_01_10_21_490500kb_count_plot <- ggplot(SAES_no_dups_01_10_21_490500kb_AG_count, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("SEVERE ADVERSE EVENTS (N)") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("ALL SAEs -> Distribution of AGE GROUPS")
SAES_no_dups_01_10_21_490500kb_count_plot
ggsave("SAES_no_dups_01_10_21_490500kb_count_plot.png")

#MERGED_01_10_21_490500kb_AGE_count <- data.frame(SAES_no_dups_01_10_21_490500kb_AG_count, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#MERGED_01_10_21_490500kb_AGE_count
```
Actual VAERS SAE percentage compared to Standard VAERS SAE percentage
```{r}
Table_SAEs <- data.frame(
  Percent_SAE = c("Standard", "Actual"),
  SAE = as.numeric(c("86000", "99838")), #the first number is the SAE/total AE
  
  stringsAsFactors = FALSE
)
Table_SAEs_perc <- Table_SAEs %>% mutate("PERC" = round(SAE/(nrow(MERGED_01_10_21_490500kb))*100))
head(Table_SAEs_perc)

Table_SAEs_perc_plot <- ggplot(Table_SAEs_perc, aes(Percent_SAE, PERC, fill = Percent_SAE)) + geom_bar(stat = 'identity') + geom_text(aes(label = PERC)) + ylab("% VAERS SAE REPORTS") + theme(axis.text.y = element_text(colour = "#000000", size = 12)) + theme(axis.text.x = element_text(angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(legend.position = "none") + ggtitle("SAEs -> Comparison of Standard versus actual % reports")
Table_SAEs_perc_plot
```
Breakthrough COVID-19 case count AND PERCENTAGE of total ID count
```{r}
strings_covid <- c("COVID-19", "Suspected COVID-19", "SARS-CoV-2 test positive", "COVID-19 pneumonia", "Asymptomatic COVID-19", "Vaccine breakthrough infection")
COVID_CASES <- function(SYMPTOM) {
  if(str_detect(SYMPTOM, paste(strings_covid, collapse = "|"))) {
      COVID19 = "COVID_CASE"
      } 
        else {
          COVID19 = "NA"
      }
    return(COVID19)
}
COVID_CASES1 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM1 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM1)) %>% filter(COVID_CASES != "NA")
#nrow(COVID_CASES1)
COVID_CASES2 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM2 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM2)) %>% filter(COVID_CASES != "NA")
#nrow(COVID_CASES2)
COVID_CASES3 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM3 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM3)) %>% filter(COVID_CASES != "NA")
#nrow(COVID_CASES3)
COVID_CASES4 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM4 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM4)) %>% filter(COVID_CASES != "NA")
#nrow(COVID_CASES4)
COVID_CASES5 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM5 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM5)) %>% filter(COVID_CASES != "NA")
#nrow(COVID_CASES5)

COVID_CASES_Total <- rbind(COVID_CASES1,COVID_CASES2,COVID_CASES3,COVID_CASES4,COVID_CASES5)
COVID_CASES_Total <- data.frame(COVID_CASES_Total, stringsAsFactors = FALSE)
#nrow(COVID_CASES_Total)
```
```{r}
COVID_CASES_Total_ID <- COVID_CASES_Total %>% distinct(VAERS_ID, .keep_all = TRUE)
#nrow(COVID_CASES_Total_ID)

COVID_CASES_Total_ENTRY <- COVID_CASES_Total_ID %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)
nrow(COVID_CASES_Total_ENTRY)
nrow(COVID_CASES_Total_ENTRY)/nrow(MERGED_01_10_21_490500kb)*100
```
```{r}
#nrow(COVID_CASES_Total)
#nrow(COVID_CASES_Total_ID)
COVID_CASES <- COVID_CASES_Total_ENTRY
#nrow(COVID_CASES)/nrow(MERGED_01_10_21_490500kb)*100
```
```{r}
AGE_GROUPS_COVID <- COVID_CASES %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
AGE_GROUPS_COVID_count <- AGE_GROUPS_COVID %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_COVID_plot <- ggplot(AGE_GROUPS_COVID_count, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF COVID-19 BREAKTHROUGH INFECTIONS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("Breakthough Covid-19 Infections -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_COVID_plot
ggsave("AGE_GROUPS_COVID_plot.png")

#AGE_GROUPS_COVID_plot_skew <- data.frame(AGE_GROUPS_COVID_count, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_COVID_plot_skew
```
```{r}
#nrow(COVID_CASES)
Table_IDs_COVID_percent <- data.frame(
  DATE = as.Date(c("1/16/21", "1/23/21", "1/30/21", "2/13/21", "2/27/21", "3/5/21", "3/12/21", "3/19/21", "3/26/21", "4/2/21", "4/9/21", "4/16/21", "4/23/21", "4/30/21", "5/7/21", "5/12/21", "5/19/21", "5/28/21", "6/4/21", "6/11/21", "6/18/21", "6/25/21", "7/2/21", "7/9/21", "7/16/21", "7/23/21", "7/30/21", "8/6/21", "8/13/21", "8/20/21", "8/27/21", "9/3/21", "9/10/21", "9/17/21", "9/24/21", "10/1/21"), "%m/%d/%y"),
  COVID = as.numeric(c("194", "254", "326", "619", "717", "776", "943", "1057", "1169", "1267", "1375", "1404", "1587", "1930", "2776", "3317", "3742", "4285", "4850", "5497", "6151", "6927", "8324", "9139", "9671", "10199", "10728", "11381", "12174", "13117", "14448", "15979", "17626", "18917", "19953", "21042")),
  
  stringsAsFactors = FALSE
)
Table_IDs_COVID_plot <- ggplot(data = Table_IDs_COVID_percent, aes(x = DATE, y = COVID)) + geom_line(data = Table_IDs_COVID_percent, aes(x = DATE, y = COVID, colour = "COVID"), colour = "red", size = 2) + geom_point(data = Table_IDs_COVID_percent, aes(DATE, COVID, colour = "COVID"), colour = "red", size = 5) + xlab("DATE") +  ylab("NUMBER OF BREAKTHROUGH INFECTIONS") + labs(title = "ABSOLUTE NUMBER OF BREAKTHROUGH INFECTIONS (CUMULATIVE)", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=15,vjust=0.5,hjust=0.5, size = 6)) + theme(panel.background = element_rect(fill = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + scale_x_continuous(breaks = round(seq(min(Table_IDs_COVID_percent$DATE), max(Table_IDs_COVID_percent$DATE), by = 10),0.1))
Table_IDs_COVID_plot
```
```{r}
COVID_CASES_count <- COVID_CASES %>% group_by(VAX_MANU, SYMPTOM1,SYMPTOM2,SYMPTOM3,SYMPTOM4,SYMPTOM5) %>% filter(VAX_MANU != "NA" & VAX_MANU != "UNKNOWN MANUFACTURER") %>% count(SYMPTOM1,SYMPTOM2,SYMPTOM3,SYMPTOM4,SYMPTOM5)
COVID_CASES_count_plot <- ggplot(COVID_CASES_count, aes(VAX_MANU, n, fill = VAX_MANU)) + geom_bar(stat = "Identity") + xlab("MANUFACTURER") + ylab("BREAKTHROUGH COVID CASES (N") + theme(axis.text.y = element_text(colour = "#000000", size = 12)) + theme(axis.text.x = element_text(angle=0,vjust=0.5,hjust=0.5, size = 10)) + ggtitle("COVID CASES AS PER MANUFACTURER") + theme(legend.position = "none")
COVID_CASES_count_plot
```
```{r}
COVID_CASES_count_VM <- COVID_CASES %>% group_by(VAX_MANU) %>% count(VAX_MANU) %>% filter(VAX_MANU != "NA" & VAX_MANU != "UNKNOWN MANUFACTURER") %>% mutate("PERC" = n/(nrow(COVID_CASES))*100) %>% mutate("Round" = round(PERC, 2))
COVID_CASES_count_VM <- data.frame(COVID_CASES_count_VM, stringsAsFactors = FALSE) 
#COVID_CASES_count_VM
```
```{r}
AGE_GROUPS_COVID_dead <- data.frame(AGE_GROUPS_COVID, stringsAsFactors = FALSE) %>% group_by(AGE_GROUP, DIED) %>% filter(DIED == "Y") %>% count(DIED) %>% mutate("SUM" = as.numeric(sum(n))) %>% mutate("Perc_D_tot" = as.numeric(n/(nrow(AGE_GROUPS_COVID))*100)) %>% mutate("Round" = round(Perc_D_tot, 2))
#AGE_GROUPS_COVID_dead
```
```{r}
no_COVID_CASES_count_VM <- COVID_CASES %>% group_by(VAX_MANU) %>% count(VAX_MANU) %>% filter(VAX_MANU != "NA" & VAX_MANU != "UNKNOWN MANUFACTURER") %>% mutate("PERC" = n/(nrow(MERGED_01_10_21_490500kb))*100) %>% mutate("Round" = round(PERC, 2))
#no_COVID_CASES_count_VM
COVID_CASES_count_VM <- COVID_CASES %>% group_by(VAX_MANU) %>% count(VAX_MANU) %>% filter(VAX_MANU != "NA" & VAX_MANU != "UNKNOWN MANUFACTURER") %>% mutate("PERC" = n/(nrow(COVID_CASES))*100) %>% mutate("Round" = round(PERC, 2))
COVID_CASES_count_VM <- data.frame(COVID_CASES_count_VM, stringsAsFactors = FALSE) 
#COVID_CASES_count_VM
```
```{r}
pie_VM <- plot_ly(COVID_CASES_count_VM, labels = ~VAX_MANU, values = ~Round, type = 'pie')
pie_VM <- pie_VM %>% layout(title = 'BREAKTHROUGH COVID-19 CASES BY MANUFACTURER',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
pie_VM
```
```{r}
pie_AGE_GROUPS_COVID_dead <- plot_ly(AGE_GROUPS_COVID_dead, labels = ~AGE_GROUP, values = ~Round, type = 'pie')
pie_AGE_GROUPS_COVID_dead <- pie_AGE_GROUPS_COVID_dead %>% layout(title = 'BREAKTHROUGH COVID-19 CASES RESULTING IN DEATH BY AGE GROUP',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
pie_AGE_GROUPS_COVID_dead
```
Female Reproductive Events count
```{r}
strings_FRI <- c("Abnormal uterine bleeding", "Afterbirth pain", "Dysmenorrhoea", "Female reproductive tract disorder", "Female reproductive tract disorder", "Alpha 1 foetoprotein abnormal", "Heavy menstrual bleeding", "Intermenstrual bleeding", "Intra-uterine contraceptive device removal", "Menstrual disorder", "Menstruation delayed", "Menstruation irregular", "Premenstrual syndrome", "Abortion spontaneous", "Abortion spontaneous incomplete", "Uterine leiomyoma", "Uterine haemorrhage", "Uterine spasm", "Bradycardia foetal", "Biopsy foetal", "Bradycardia foetal", "Foetal cystic hygroma", "Foetal death", "Foetal exposure during pregnancy", "Foetal heart rate abnormal", "Foetal hypokinesia", "Foetal non-stress test abnormal", "Amenorrhoea", "Hypomenorrhoea", "Menopause", "Menorrhagia", "Oligomenorrhoea", "Polymenorrhoea", "Postmenopausal haemorrhage", "Postmenopause", "Vaginal discharge", "Vaginal disorder", "Vaginal erosion", "Vaginal haemorrhage", "Vaginal lesion", "Vaginal infection", "Vulvovaginal mycotic infection", "Vulval ulceration", "Vulval abscess", "Vulva cyst", "Cervix inflammation", "Adenocarcinoma of the cervix", "Cervix carcinoma", "Amniorrhoea", "Amniotic cavity infection", "Amniotic membrane rupture test positive", "Amniotic fluid index decreased")
strings_FRI
```
```{r}
FRI <- function(SYMPTOM) {
  if(str_detect(SYMPTOM, paste(strings_FRI, collapse = "|"))) {
      FRI = "Female_Reprod_Issues"
      } 
        else {
          FRI = "NA"
      }
    return(FRI)
}
MERGED_01_10_21_490500kb_FRI1 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM1 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM1)) %>% filter(Female_Reprod_Issues != "NA")
MERGED_01_10_21_490500kb_FRI2 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM2 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM2)) %>% filter(Female_Reprod_Issues != "NA")
MERGED_01_10_21_490500kb_FRI3 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM3 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM3)) %>% filter(Female_Reprod_Issues != "NA")
MERGED_01_10_21_490500kb_FRI4 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM4 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM4)) %>% filter(Female_Reprod_Issues != "NA")
MERGED_01_10_21_490500kb_FRI5 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM5 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM5)) %>% filter(Female_Reprod_Issues != "NA")

MERGED_01_10_21_490500kb_FRI_Total <- rbind(MERGED_01_10_21_490500kb_FRI1,MERGED_01_10_21_490500kb_FRI2,MERGED_01_10_21_490500kb_FRI3,MERGED_01_10_21_490500kb_FRI4,MERGED_01_10_21_490500kb_FRI5)
MERGED_01_10_21_490500kb_FRI_Total <- data.frame(MERGED_01_10_21_490500kb_FRI_Total, stringsAsFactors = FALSE)

MERGED_01_10_21_490500kb_FRI_Total_ID <- MERGED_01_10_21_490500kb_FRI_Total %>% distinct(VAERS_ID, .keep_all = TRUE)

MERGED_01_10_21_490500kb_FRI_Total_ENTRY <- MERGED_01_10_21_490500kb_FRI_Total_ID %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)

#nrow(MERGED_01_10_21_490500kb_FRI_Total_ENTRY)
MERGED_01_10_21_490500kb_FRI_Total_ENTRY <- data.frame(MERGED_01_10_21_490500kb_FRI_Total_ENTRY, stringsAsFactors = FALSE)
```

```{r}
FRI_CASES <- MERGED_01_10_21_490500kb_FRI_Total_ENTRY
nrow(FRI_CASES)
nrow(FRI_CASES)/nrow(MERGED_01_10_21_490500kb)*100
```
Histogram showing distribution of Female Reproductive Issues by age group
```{r}
FRI_CASES_ID <- FRI_CASES %>% group_by(AGE_YRS) %>% filter(AGE_YRS != "NA") %>% count(AGE_YRS)
FRI_CASES_ID_plot <- ggplot(FRI_CASES_ID, aes(AGE_YRS, n, fill = AGE_YRS)) + geom_bar(stat = "Identity") + xlab("AGE YEARS") + ylab("Number of Women with Female Reproductive Issues") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14))  + theme(legend.position = "none") + ggtitle("Number of Female Reproductive Issues -> VAERS IDs vs AGE YEARS") + xlim(0,100)
FRI_CASES_ID_plot
ggsave("FRI_CASES_ID_plot.png")
```
```{r}
FRI_CASES$ONSET_DATE <- as.Date(FRI_CASES$ONSET_DATE, "%m/%d/%y")
FRI_CASES$VAX_DATE <- as.Date(FRI_CASES$VAX_DATE, "%m/%d/%y")


FRI_CASES <- data.frame(FRI_CASES, stringsAsFactors = FALSE) %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2020) 
FRI_CASES_VD_OD <- FRI_CASES %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) %>% arrange(desc(n)) %>% mutate("OBSERVED" = n) %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = as.integer(mean(n))) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(FRI_CASES)*100)
#head(FRI_CASES_VD_OD)

chisq.test(FRI_CASES_VD_OD$OBSERVED)
```
Cumulative FRI cases as per update - rates increasing. 
```{r}
Table_IDs_percent_plot_FRI <- ggplot() + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, FRI), colour = "red", size = 2) + geom_point(data = Table_IDs_percent_FV_IDs, aes(DATE, FRI), colour = "red", colour = "red", size = 5) + labs(caption = "Data source: VAERS; Analysis: Dr. Jessica Rose") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + theme(legend.background = element_rect(fill = "white"), legend.key = element_rect(fill = "black", color = NA), legend.key.size = unit(0.5, "cm"), legend.key.width = unit(0.5,"cm")) 
Table_IDs_percent_plot_FRI
ggsave("Table_IDs_percent_plot_FRI.png")
```
Heatmap showing Female Reproductive Issues wrt VAX DATE and ONSET DATE. Early April shows a clustering of intersecting points. 
```{r}
FRI_CASES_VD_OD_TILE_count <- data.frame(FRI_CASES, stringsAsFactors = FALSE) %>% group_by(VAX_DATE, ONSET_DATE) %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021) %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=350) %>% count(diff_in_days) %>% arrange(desc(diff_in_days))  

FRI_CASES_VD_OD_TILE_plot <- ggplot(FRI_CASES_VD_OD_TILE_count, aes(VAX_DATE, ONSET_DATE)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("ONSET OF AEs") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.x = element_text(colour = "#000000", size = 14)) + theme(axis.text.y = element_text(colour = "#000000", size = 14), axis.text.x=element_text(angle=0,vjust=0.5,hjust=0.5)) + ggtitle("VACCINE DATE vs FRI AE ONSET DATE") 
FRI_CASES_VD_OD_TILE_plot 
ggsave("FRI_CASES_VD_OD_TILE_plot.png")

#FRI_CASES_VD_OD_TILE_count
```

```{r}
CAUSATION_FRI_CASES_plot <- ggplot() + geom_line(data = FRI_CASES_VD_OD, aes(diff_in_days, Percentage_OBSERVED), colour = "orange", size = 2) + ylab("PERCENTAGE REPORTED FRI AEs") + xlab("NUMBER OF DAYS FOLLOWING INJECTION") + geom_line(data = FRI_CASES_VD_OD, aes(diff_in_days, Percentage_EXPECTED), colour = "yellow", size = 2) + labs(title = "FEMALE REPRODUCTIVE ISSUES post injection", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
CAUSATION_FRI_CASES_plot 
ggsave("CAUSATION_FRI_CASES_plot.png")
```
Cardiac Events Count
```{r}
strings_CV <- c("Arrhythmia", "Myocarditis", "Pericarditis", "Endocarditis", "Cardiac arrest", "Chest pain", "Myocardial infarction", "Pericardial effusion", "Aneurysm", "Blood fibrinogen", "Blood fibrinogen increased", "Circulatory collapse", "C-reactive protein decreased", "C-reactive protein increased", "Deep vein thrombosis", "Dizziness", "Fatigue", "Fibrin D dimer increased", "Irregular breathing", "Ischaemia", "Microembolism", "Pallor", "Palpitations", "Platelet count decreased", "Platelet count increased", "Red blood cell count abnormal", "Red blood cell count decreased", "Red blood cell count increased", "Red blood cell rouleaux formation present", "Red blood cell schistocytes present", "Syncope", "Troponin", "Troponin increased", "Troponin I", "Troponin I increased", "Troponin T increased", "Thrombosis", "Echocardiogram abnormal", "Dyspnoea")
strings_CV
Cardiack <- function(SYMPTOM) {
  if(str_detect(SYMPTOM, paste(strings_CV, collapse = "|"))) {
      CV = "Cardiac_event"
      } 
        else {
          CV = "NA"
      }
    return(CV)
}
Cardiack1 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM1 != "NA") %>% mutate("Cardiack" = Cardiack(SYMPTOM1)) %>% filter(Cardiack != "NA")
#nrow(Cardiack1)
Cardiack2 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM2 != "NA") %>% mutate("Cardiack" = Cardiack(SYMPTOM2)) %>% filter(Cardiack != "NA")
#nrow(Cardiack2)
Cardiack3 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM3 != "NA") %>% mutate("Cardiack" = Cardiack(SYMPTOM3)) %>% filter(Cardiack != "NA")
#nrow(Cardiack3)
Cardiack4 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM4 != "NA") %>% mutate("Cardiack" = Cardiack(SYMPTOM4)) %>% filter(Cardiack != "NA")
#nrow(Cardiack4)
Cardiack5 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM5 != "NA") %>% mutate("Cardiack" = Cardiack(SYMPTOM5)) %>% filter(Cardiack != "NA")
#nrow(Cardiack5)

Cardiack_Total <- rbind(Cardiack1,Cardiack2,Cardiack3,Cardiack4,Cardiack5)
Cardiack_Total <- data.frame(Cardiack_Total, stringsAsFactors = FALSE)
#nrow(Cardiack_Total)
```
```{r}
Cardiack_Total_ID <- Cardiack_Total %>% distinct(VAERS_ID, .keep_all = TRUE)
#nrow(Cardiack_Total_ID)

Cardiack_Total_ENTRY <- Cardiack_Total_ID %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)
nrow(Cardiack_Total_ENTRY)
```
```{r}
#nrow(Cardiack_Total)
#nrow(Cardiack_Total_ID)
Cardiack <- Cardiack_Total_ENTRY
#nrow(Cardiack)/nrow(MERGED_01_10_21_490500kb)*100
```
```{r}
AGE_GROUPS_Cardiack <- Cardiack %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
AGE_GROUPS_Cardiack_count <- AGE_GROUPS_Cardiack %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_Cardiack_plot <- ggplot(AGE_GROUPS_Cardiack_count, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF CARDIAC EVENTS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("CARDIAC EVENTS -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_Cardiack_plot
ggsave("AGE_GROUPS_Cardiack_plot.png")

#AGE_GROUPS_Cardiack_plot_skew <- data.frame(AGE_GROUPS_Cardiack_count, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_Cardiack_plot_skew
```
This is the number of kids aged 12-18 reported to VAERS with AEs, the percentage of the total AEs reported and a histogram showing age group distribution.
```{r}
MERGED_01_10_21_490500kb_12_18 <- MERGED_01_10_21_490500kb %>% filter(AGE_YRS >=12 & AGE_YRS <=18)
nrow(MERGED_01_10_21_490500kb_12_18)

MERGED_01_10_21_490500kb_12_18_1 <- MERGED_01_10_21_490500kb_12_18 %>% distinct(VAERS_ID, .keep_all = TRUE)

YOUTH_CASES <- MERGED_01_10_21_490500kb_12_18_1 %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)

YOUTH_CASES_ID <- YOUTH_CASES %>% group_by(AGE_YRS) %>% filter(AGE_YRS != "NA") %>% count(AGE_YRS)
YOUTH_CASES_ID_plot <- ggplot(YOUTH_CASES_ID, aes(AGE_YRS, n, fill = AGE_YRS)) + geom_bar(stat = "Identity") + xlab("AGE YEARS") + ylab("Number of YOUTH 12-18 y/o CASES") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14))  + theme(legend.position = "none") + ggtitle("Number of AEs in 12-18 y/os  -> VAERS IDs vs AGE YEARS") #+ xlim(0,100)
YOUTH_CASES_ID_plot
ggsave("YOUTH_CASES_ID_plot.png")
```
```{r}
YOUTH_CASES <- data.frame(YOUTH_CASES, stringsAsFactors = FALSE) %>% group_by(VAX_DATE, ONSET_DATE) %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2021)

YOUTH_CASES_VD_OD <- YOUTH_CASES %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=150) %>% count(diff_in_days) %>% arrange(desc(n)) %>% mutate("OBSERVED" = n) %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = as.integer(mean(n))) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(YOUTH_CASES)*100)
#head(YOUTH_CASES_VD_OD)

#chisq.test(YOUTH_CASES_VD_OD$OBSERVED)
nrow(YOUTH_CASES)
nrow(YOUTH_CASES)/nrow(MERGED_01_10_21_490500kb)*100
```
And this is a heatmap showing 1:1 correlation between onset of AE and injection date.
```{r}
YOUTH_CASES_VD_OD_TILE_count <- data.frame(YOUTH_CASES, stringsAsFactors = FALSE) %>% group_by(VAX_DATE, ONSET_DATE) %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021) %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=350) %>% count(diff_in_days) %>% arrange(desc(n))  

YOUTH_CASES_VD_DD_TILE_plot <- ggplot(YOUTH_CASES_VD_OD_TILE_count, aes(VAX_DATE, ONSET_DATE)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("ONSET OF AEs") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.x = element_text(colour = "#000000", size = 14)) + theme(axis.text.y = element_text(colour = "#000000", size = 14), axis.text.x=element_text(angle=0,vjust=0.5,hjust=0.5)) + ggtitle("VACCINE DATE vs YOUTH CASES 12-15 y/o AE ONSET DATE") 
YOUTH_CASES_VD_DD_TILE_plot 
ggsave("FYOUTH_CASES_VD_DD_TILE_plot.png")

YOUTH_CASES_VD_OD_TILE_count
```
Here we have the percentage of cardiac cases for kids aged 12-15 of the total youth population (12-18) in VAERS, the percentage of cardiac AEs that come from youths and the percentage of cardiac cases in kids aged 12-15 of the total cardiac AEs. 
```{r}
Cardiack_12_15 <- Cardiack %>% filter(AGE_YRS >=12 & AGE_YRS <=15)

nrow(Cardiack_12_15)/nrow(YOUTH_CASES)*100
nrow(YOUTH_CASES)/nrow(Cardiack)*100
nrow(Cardiack_12_15)/nrow(Cardiack)*100
```
Immunological Adverse Effects Count
```{r}
strings_IAE <- c("Anaphylactic reaction", "Anaphylactic shock", "Anaphylactoid reaction", "Autoimmune demyelinating disease", "Autoimmune disorder", "Autoimmune thyroiditis", "Chills", "Cytokine storm", "Drug reaction with eosinophilia and systemic symptoms", "Dysgeusia", "Dysphagia", "Encephalomyelitis", "Erythema", "Febrile neutropenia", "Fungal infection", "Genital rash",  "Guillain-Barre syndrome", "Haemorrhage subepidermal", "Hepatitis", "Hepatitis acute", "Herpes simplex", "Herpes simplex encephalitis", "Herpes simplex reactivation", "Herpes virus infection", "Herpes zoster", "Herpes zoster cutaneous disseminated", "Hyperpyrexia", "Immediate post-injection reaction", "Immune thrombocytopenia", "Inflammation", "Leukopenia", "Lymphadenitis", "Lymphadenopathy", "Malaise", "May-Thurner syndrome", "Meningitis", "Meningitis aseptic", "Meningitis viral", "Myelitis", "Multiple sclerosis relapse", "Multiple allergies", "Myelitis transverse", "Neuritis", "Noninfective encephalitis", "Ophthalmic herpes simplex", "Oral herpes", "Oral viral infection", "Pancreatitis", "Pancreatitis acute", "Parkinsonism", "Peptostreptococcus infection", "Periarthritis", "Pneumonia viral", "Pruritus", "Rash", "Rash erythematous", "Rash macular", "Rash maculo-papular", "Rash morbilliform", "Rash papular", "Rash pruritic", "Rash pustular", "Rash vesicular", "Reaction to excipient", "Rheumatoid arthritis", "Butterfly rash", "Necrosis", "Stevens-Johnson syndrome", "Systemic lupus erythematosus", "Systemic lupus erythematosus rash", "Systemic scleroderma", "Thrombophlebitis", "Type III immune complex mediated reaction", "Urticaria", "Varicella zoster virus infection", "Vestibular neuronitis", "Viral cardiomyopathy", "Viral infection", "Viral pericarditis", "Viral rash", "Systemic scleroderma")
strings_IAE
IAE <- function(SYMPTOM) {
  if(str_detect(SYMPTOM, paste(strings_IAE, collapse = "|"))) {
      IAE = "Immunological_Effect"
      } 
        else {
          IAE = "NA"
      }
    return(IAE)
}
IAE1 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM1 != "NA") %>% mutate("IAE" = IAE(SYMPTOM1)) %>% filter(IAE != "NA")
#nrow(IAE1)
IAE2 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM2 != "NA") %>% mutate("IAE" = IAE(SYMPTOM2)) %>% filter(IAE != "NA")
#nrow(IAE2)
IAE3 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM3 != "NA") %>% mutate("IAE" = IAE(SYMPTOM3)) %>% filter(IAE != "NA")
#nrow(IAE3)
IAE4 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM4 != "NA") %>% mutate("IAE" = IAE(SYMPTOM4)) %>% filter(IAE != "NA")
#nrow(IAE4)
IAE5 <- MERGED_01_10_21_490500kb %>% rowwise() %>% filter(SYMPTOM5 != "NA") %>% mutate("IAE" = IAE(SYMPTOM5)) %>% filter(IAE != "NA")
#nrow(IAE5)

IAE_Total <- rbind(IAE1,IAE2,IAE3,IAE4,IAE5)
IAE_Total <- data.frame(IAE_Total, stringsAsFactors = FALSE)
#nrow(IAE_Total)
```
```{r}
IAE_Total_ID <- IAE_Total %>% distinct(VAERS_ID, .keep_all = TRUE)
#nrow(IAE_Total_ID)
```
Here we have the total AE count for immunological adverse events as per the list above. This is a very short list as compared with the total number of immunological AEs. It is merely a representation. 
```{r}
IAE_Total_ENTRY <- IAE_Total_ID %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)
nrow(IAE_Total_ENTRY)
```
```{r}
#nrow(IAE_Total)
#nrow(IAE_Total_ID)
IAE_CASES <- IAE_Total_ENTRY
#nrow(IAE)/nrow(MERGED_01_10_21_490500kb)*100
```
```{r}
AGE_GROUPS_IAE <- IAE_CASES %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
AGE_GROUPS_IAE_count <- AGE_GROUPS_IAE %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_IAE_plot <- ggplot(AGE_GROUPS_IAE_count, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF IMMUNOLOGICAL ADVERSE EVENTS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("Immunological Adverse Event -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_IAE_plot
ggsave("AGE_GROUPS_IAE_plot.png")

#AGE_GROUPS_IAE_plot_skew <- data.frame(AGE_GROUPS_IAE_count, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_Cardiack_plot_skew
```
This is a histogram showing death distribution according to age group. We also have a time series plot showing clustering of data around day 1 following injection. And! We have a heatmap confirming strong correlation between injection date and death date. R=1 -> perfect correlation (look for red on diagonal).
```{r}
D <- MERGED_01_10_21_490500kb %>% filter(DIED == "Y") %>% filter(year(DATEDIED) <= 2021 & year(DATEDIED) >= 2020) 
MERGED_01_10_21_490500kb_VD_DD <- D %>% mutate("diff_in_days" = difftime(DATEDIED, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) %>% arrange(desc(n)) %>% mutate("OBSERVED" = n) %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = as.integer(mean(n))) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(D)*100)
#head(MERGED_01_10_21_490500kb_VD_DD)

#chisq.test(MERGED_01_10_21_490500kb_VD_DD$OBSERVED)

H <- MERGED_01_10_21_490500kb %>% filter(HOSPITAL == "Y") %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2020)
MERGED_01_10_21_490500kb_H_VD_OD <- H %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) %>% arrange(desc(n))  %>% mutate("OBSERVED" = n)  %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = mean(n)) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(H)*100)
#head(MERGED_01_10_21_490500kb_H_VD_OD)

#chisq.test(MERGED_01_10_21_490500kb_H_VD_OD$OBSERVED)

ER <- MERGED_01_10_21_490500kb %>% filter(ER_ED_VISIT == "Y") %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2020)
MERGED_01_10_21_490500kb_ER_VD_OD <- ER %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) %>% arrange(desc(n))  %>% mutate("OBSERVED" = n) %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = mean(n)) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(ER)*100)
#head(MERGED_01_10_21_490500kb_ER_VD_OD)

#chisq.test(MERGED_01_10_21_490500kb_ER_VD_OD$OBSERVED)


AGE_GROUPS_D <- D %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
AGE_GROUPS_H <- H %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
AGE_GROUPS_ER <- ER %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
```
Here is a histogram showing death distribution by age group.
```{r}
AGE_GROUPS_D_ALL_ID <- AGE_GROUPS_D %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_D_ALL_ID_plot <- ggplot(AGE_GROUPS_D_ALL_ID, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF DEATHS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("Death -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_D_ALL_ID_plot
ggsave("AGE_GROUPS_D_ALL_ID_plot.png")
#AGE_GROUPS_D_ALL_ID_skew <- data.frame(AGE_GROUPS_D_ALL_ID, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_D_ALL_ID_skew
```
Here is a plot showing clustering of reports around 0 and 1 in the context of death. 
```{r}
CAUSATION_DEATH_plot <- ggplot() + geom_line(data = MERGED_01_10_21_490500kb_VD_DD, aes(diff_in_days, Percentage_OBSERVED), colour = "orange", size = 2) + ylab("PERCENTAGE OF REPORTED DEATHS") + xlab("NUMBER OF DAYS FOLLOWING INJECTION") + geom_line(data = MERGED_01_10_21_490500kb_VD_DD, aes(diff_in_days, Percentage_EXPECTED), colour = "yellow", size = 2) + labs(title = "DEATHS post injection", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(panel.background = element_rect(fill = "#100000", color = "pink")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
CAUSATION_DEATH_plot 
ggsave("CAUSATION_DEATH_plot.png")
```
```{r}
VD_DD_TILE_count <- MERGED_01_10_21_490500kb %>% group_by(DIED, DATEDIED, VAX_DATE) %>% filter(DIED == "Y") %>% filter(year(DATEDIED) <= 2021 & year(DATEDIED) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021) %>% mutate("diff_in_days" = difftime(DATEDIED, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=350) %>% count(diff_in_days) 

#png("VD_DD_TILE_plot ", width = 4, height = 4, units = 'in', res = 100)

VD_DD_TILE_plot <- ggplot(VD_DD_TILE_count, aes(VAX_DATE, DATEDIED)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("DATE OF DEATH") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + ggtitle("VACCINE DATE vs DEATH DATE OF AEs") 
VD_DD_TILE_plot 
ggsave("VD_DD_TILE_plot.png")
```
This is a histogram showing hospitalization distribution according to age group. We also have a time series plot showing clustering of data around day 1 following injection. And! We have a heatmap confirming strong correlation between injection date and hospitalization date. R=1 -> perfect correlation (look for red on diagonal).
```{r}
AGE_GROUPS_H_ALL_ID <- AGE_GROUPS_H %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_H_ALL_ID_plot <- ggplot(AGE_GROUPS_H_ALL_ID, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF HOSPITALIZATIONS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("Hospital -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_H_ALL_ID_plot
ggsave("AGE_GROUPS_H_ALL_ID_plot.png")
#AGE_GROUPS_H_ALL_ID_skew <- data.frame(AGE_GROUPS_H_ALL_ID, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_H_ALL_ID_skew
```

```{r}
CAUSATION_HOSP_plot <- ggplot() + geom_line(data = MERGED_01_10_21_490500kb_H_VD_OD, aes(diff_in_days, Percentage_OBSERVED), colour = "orange", size = 2) + xlab("NUMBER OF DAYS FOLLOWING INJECTION") +  ylab("PERCENTAGE OF REPORTED HOSPITALIZATIONS") + geom_line(data = MERGED_01_10_21_490500kb_H_VD_OD, aes(diff_in_days, Percentage_EXPECTED), colour = "yellow", size = 2) + labs(title = "HOSPITALIZATIONS post injection", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000", color = "pink")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
CAUSATION_HOSP_plot
ggsave("CAUSATION_HOSP_plot.png")
```

```{r}
H_VD_OD_TILE_count <- MERGED_01_10_21_490500kb %>% group_by(HOSPITAL, ONSET_DATE, VAX_DATE) %>% filter(HOSPITAL == "Y") %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021)%>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=350) %>% count(diff_in_days) 

H_VD_DD_TILE_plot <- ggplot(H_VD_OD_TILE_count, aes(VAX_DATE, ONSET_DATE)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("ONSET OF AEs") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.x = element_text(colour = "#000000", size = 14)) + theme(axis.text.y = element_text(colour = "#000000", size = 14), axis.text.x=element_text(angle=0,vjust=0.5,hjust=0.5)) + ggtitle("VACCINE DATE vs HOSPITALIZATION DATE") 
H_VD_DD_TILE_plot 
ggsave("H_VD_DD_TILE_plot.png")
```
This is a histogram showing ER distribution according to age group. We also have a time series plot showing clustering of data around day 1 following injection. And! We have a heatmap confirming strong correlation between injection date and ER date. R=1 -> perfect correlation (look for red on diagonal).
```{r}
AGE_GROUPS_ER_ALL_ID <- AGE_GROUPS_ER %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_ER_ALL_ID_plot <- ggplot(AGE_GROUPS_ER_ALL_ID, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF EMERGENCY DOCTOR VISITS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("ER -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_ER_ALL_ID_plot
ggsave("AGE_GROUPS_ER_ALL_ID_plot.png")
#AGE_GROUPS_ER_ALL_ID_skew <- data.frame(AGE_GROUPS_ER_ALL_ID, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_ER_ALL_ID_skew
```

```{r}
CAUSATION_ER_plot <- ggplot() + geom_line(data = MERGED_01_10_21_490500kb_ER_VD_OD, aes(diff_in_days, Percentage_OBSERVED), colour = "orange", size = 2) + xlab("NUMBER OF DAYS FOLLOWING INJECTION") +  ylab("PERCENTAGE OF REPORTED ER VISITS") + geom_line(data = MERGED_01_10_21_490500kb_ER_VD_OD, aes(diff_in_days, Percentage_EXPECTED), colour = "yellow", size = 2) + labs(title = "ER VISITS post injection", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000", color = "pink")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
CAUSATION_ER_plot 
ggsave("CAUSATION_ER_plot.png")
```

```{r}
ER_VD_OD_TILE_count <- MERGED_01_10_21_490500kb %>% group_by(ER_ED_VISIT, ONSET_DATE, VAX_DATE) %>% filter(ER_ED_VISIT == "Y") %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021)%>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=350) %>% count(diff_in_days) 

ER_VD_DD_TILE_plot <- ggplot(ER_VD_OD_TILE_count, aes(VAX_DATE, ONSET_DATE)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("ONSET OF AEs") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.x = element_text(colour = "#000000", size = 14)) + theme(axis.text.y = element_text(colour = "#000000", size = 14), axis.text.x=element_text(angle=0,vjust=0.5,hjust=0.5)) + ggtitle("VACCINE DATE vs ER visit DATE") 
ER_VD_DD_TILE_plot 
ggsave("ER_VD_DD_TILE_plot.png")
```
And hey! What would this analysis be without myocarditis? This is the count of myocarditis cases and the percentage of the total count for cardiac cases and for the total AE count. 
```{r}
MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo1 <- MERGED_01_10_21_490500kb %>% filter(SYMPTOM1 == "Myocarditis") 
#nrow(MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo1)
MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo2 <- MERGED_01_10_21_490500kb %>% filter(SYMPTOM2 == "Myocarditis") 
#nrow(MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo2)
MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo3 <- MERGED_01_10_21_490500kb %>% filter(SYMPTOM3 == "Myocarditis") 
#nrow(MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo3)
MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo4 <- MERGED_01_10_21_490500kb %>% filter(SYMPTOM4 == "Myocarditis")
#nrow(MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo4)
MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo5 <- MERGED_01_10_21_490500kb %>% filter(SYMPTOM5 == "Myocarditis") 
#nrow(MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo5)
myo_Total <- rbind(MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo1,MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo2,MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo3,MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo4,MERGED_SYM_DAT_VAX_01_10_21_490500kb_myo5)
myo_Total <- data.frame(myo_Total, stringsAsFactors = FALSE)
nrow(myo_Total)
nrow(myo_Total)/nrow(Cardiack)*100
nrow(myo_Total)/nrow(MERGED_01_10_21_490500kb)*100
#myo_Total
myo_Total_count <- myo_Total %>% count(SYMPTOM1,SYMPTOM2,SYMPTOM3,SYMPTOM4,SYMPTOM5) %>% mutate("SUM" = sum(n)) %>% arrange(desc(n))
#myo_Total_count

myo_Total_ID <- myo_Total %>% distinct(VAERS_ID, .keep_all = TRUE)

myo_Total_ENTRY <- myo_Total_ID %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)

#nrow(myo_Total)
#nrow(myo_Total_ENTRY)

myo_Total_doses_all <- myo_Total  %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n))
#myo_Total_doses_all

myo_Total_doses_1_2 <- myo_Total  %>% filter(VAX_DOSE_SERIES <= 2) %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n))
#myo_Total_doses_1_2

myo_Total_doses_1_2_age <- myo_Total  %>% filter(VAX_DOSE_SERIES <= 2) %>% group_by(AGE_YRS) %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n))
#myo_Total_doses_1_2_age 
```
And this plot shows age versus dose data for myocarditis cases. Notice how many young people. Hmm what's that now? Dose 2 is causing moycarditis in young males? Hmm
```{r}
AGE_YEARS_myo_Total_plot <- ggplot(myo_Total_doses_1_2_age, aes(AGE_YRS, n, fill = VAX_DOSE_SERIES)) + geom_bar(stat = "identity", position = position_dodge(preserve = 'single')) + xlab("Age of reported patient, years") + ylab("Absolute number of Myocarditis") + labs(caption = "Data source: VAERS/Analysis: Dr. Jessica Rose") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(legend.position = "right") + ggtitle("Myocarditis in VAERS after mRNA injection by age and dose #") + theme(legend.text=element_text(size=6), legend.title=element_text(size=6)) + xlim(12,62) + geom_vline(xintercept=c(14.5,16)) #+ ylim(0,20)  
AGE_YEARS_myo_Total_plot
ggsave("AGE_YEARS_myo_Total_plot.png")
```
```{r}
myo_Total_ENTRY_VDS <- myo_Total_ENTRY %>% filter(VAX_DOSE_SERIES != "NA" & VAX_DOSE_SERIES != "UNK") %>% filter(SEX != "U") %>% filter(VAX_MANU != "JANSSEN")
myo_Total_ENTRY_count_VDS_plot <- ggplot(myo_Total_ENTRY_VDS, aes(VAX_DOSE_SERIES, AGE_YRS, fill = SEX)) + geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=1) + facet_wrap(~ VAX_MANU, scale = "free") + xlab("DOSE NUMBER") + ylab("AGE IN YEARS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "right") + ggtitle("MYOCARDITIS -> VAERS IDs vs AGE_YRS") 
myo_Total_ENTRY_count_VDS_plot
```