---
title: "The R Markdown document for VAERS analysis"
author: "Dr. Jessica Rose"
date: "2021-05-21"
output: html_document
---

This is an R Markdown sheet generated from the VAERS data. I included some code but for the most part, only figures are shown. I decided to show figures pertaining to Death, Spontaneous abortions, Breakthrough COVID-19 infections and Cardiovascular, Neurological and Immunological adverse events.

I start by coalescing the .csv files downloaded from the VAERS website: https://vaers.hhs.gov/data/datasets.html

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=5, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```
```{r}
library("readr")
library("tidyverse")
library("dplyr")
library("ggplot2")
library("stringr")
library("DT")
library("reshape2")
library("data.table")
library("plotly")
```
```{r}
VAERSSYM_07_02_21_418000kb <- read_csv("/Users/admin/Downloads/2021VAERSData_07_02_21_418000kb/2021VAERSSYMPTOMS.csv", col_types = cols(VAERS_ID = col_double()))
VAERSDATA_07_02_21_418000kb <- read_csv("/Users/admin/Downloads/2021VAERSData_07_02_21_418000kb/2021VAERSDATA.csv", col_types = cols(VAERS_ID = col_double()))
VAERSVAX_07_02_21_418000kb <- read_csv("/Users/admin/Downloads/2021VAERSData_07_02_21_418000kb/2021VAERSVAX.csv", col_types = cols(VAERS_ID = col_double()))
RV_VAX_07_02_21_418000kb <- VAERSVAX_07_02_21_418000kb %>% filter(VAX_TYPE == "COVID19")
MERGED_SYM_DAT_07_02_21_418000kb <- full_join(VAERSSYM_07_02_21_418000kb, VAERSDATA_07_02_21_418000kb, by = c("VAERS_ID"))
MERGED_SYM_DAT_VAX_07_02_21_418000kb <- full_join(MERGED_SYM_DAT_07_02_21_418000kb, RV_VAX_07_02_21_418000kb, by = c("VAERS_ID"))
MERGED_07_02_21_418000kb_1 <- MERGED_SYM_DAT_VAX_07_02_21_418000kb %>% distinct(VAERS_ID, .keep_all = TRUE)
MERGED_07_02_21_418000kb <- MERGED_07_02_21_418000kb_1 %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)

write.csv(MERGED_07_02_21_418000kb, "/Users/admin/Desktop/MERGED_07_02_21_418000kb.csv")
```
```{r}
MERGED_07_02_21_418000kb$VAX_DATE <- as.Date(MERGED_07_02_21_418000kb$VAX_DATE, "%m/%d/%y") 
MERGED_07_02_21_418000kb$ONSET_DATE <- as.Date(MERGED_07_02_21_418000kb$ONSET_DATE, "%m/%d/%y")
MERGED_07_02_21_418000kb$DATEDIED <- as.Date(MERGED_07_02_21_418000kb$DATEDIED, "%m/%d/%y")
MERGED_07_02_21_418000kb$TODAYS_DATE <- as.Date(MERGED_07_02_21_418000kb$TODAYS_DATE, "%m/%d/%y")
MERGED_07_02_21_418000kb$RECVDATE <- as.Date(MERGED_07_02_21_418000kb$RECVDATE, "%m/%d/%y")

nrow(MERGED_07_02_21_418000kb)
```
```{r}
AGE_GROUPED_BY_DECADE <- function(AGE_YRS) {
  if(AGE_YRS >= 0 & AGE_YRS <10) {
    AGE_GROUP = "A. 0-10"
    } else if(AGE_YRS >= 10 & AGE_YRS <20) {
      AGE_GROUP = "B. 10-20"
      } else if(AGE_YRS >= 20 & AGE_YRS <30) {
        AGE_GROUP = "C. 20-30"
        } else if(AGE_YRS >= 30 & AGE_YRS <40) {
          AGE_GROUP = "D. 30-40"
          } else if(AGE_YRS >= 40 & AGE_YRS <50) {
            AGE_GROUP = "E. 40-50"
            } else if(AGE_YRS >= 50 & AGE_YRS <60) {
              AGE_GROUP = "F. 50-60"
              } else if(AGE_YRS >= 60 & AGE_YRS <70) {
                AGE_GROUP = "G. 60-70"
                } else if(AGE_YRS >= 70 & AGE_YRS <80) {
                  AGE_GROUP = "H.70-80"
                  } else if(AGE_YRS >= 80 & AGE_YRS <90) {
                    AGE_GROUP = "I. 80-90"
                    } else if(AGE_YRS >= 90 & AGE_YRS <100) {
                      AGE_GROUP = "J. 90-100"
                      } else if(AGE_YRS >= 100 & AGE_YRS <110) {
                        AGE_GROUP = "K. 100-110"
                        }
  else {
    AGE_GROUP = "NA"
    }
  return(AGE_GROUP)
}
```
```{r}
MERGED_07_02_21_418000kb_AG <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
MERGED_07_02_21_418000kb_AG_count <- MERGED_07_02_21_418000kb_AG %>% count(AGE_GROUP)
MERGED_07_02_21_418000kb_count_plot <- ggplot(MERGED_07_02_21_418000kb_AG_count, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("ADVERSE EVENTS (N)") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("ALL AEs -> Distribution of AGE GROUPS")
MERGED_07_02_21_418000kb_count_plot
ggsave("MERGED_07_02_21_418000kb_count_plot.png")

MERGED_07_02_21_418000kb_AGE_count <- MERGED_07_02_21_418000kb %>% group_by(AGE_YRS) %>% filter(AGE_YRS != "NA") %>% count(AGE_YRS)
#MERGED_07_02_21_418000kb_AGE_count <- data.frame(MERGED_07_02_21_418000kb_AGE_count, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#MERGED_07_02_21_418000kb_AGE_count
```
```{r}
MERGED_07_02_21_418000kb_doses <- MERGED_07_02_21_418000kb %>% group_by(VAX_DOSE_SERIES) %>% filter(VAX_DOSE_SERIES >= 0 & VAX_DOSE_SERIES <= 2) %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n))
#MERGED_07_02_21_418000kb_doses
MERGED_07_02_21_418000kb_doses_VM <- MERGED_07_02_21_418000kb %>% group_by(VAX_DOSE_SERIES, VAX_MANU) %>% filter(VAX_DOSE_SERIES >= 0 & VAX_DOSE_SERIES <= 2 & VAX_MANU != "UNKNOWN MANUFACTURER") %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n)) 
MERGED_07_02_21_418000kb_doses_n <- MERGED_07_02_21_418000kb_doses %>% mutate("SUM" = sum(MERGED_07_02_21_418000kb_doses$n)) %>% mutate("PERCENT" = n/SUM*100)
#MERGED_07_02_21_418000kb_doses_n
```
```{r}
BD_07_02_21_418000kb <- MERGED_07_02_21_418000kb %>% filter(BIRTH_DEFECT == 1)
nrow(BD_07_02_21_418000kb)
```
```{r}
D_07_02_21_418000kb <- MERGED_07_02_21_418000kb %>% group_by(VAERS_ID, DIED) %>% filter(DIED == "Y")
nrow(D_07_02_21_418000kb)
```
```{r}
HOSPITAL_07_02_21_418000kb  <- MERGED_07_02_21_418000kb  %>% group_by(VAERS_ID, HOSPITAL) %>% filter(HOSPITAL == "Y")
nrow(HOSPITAL_07_02_21_418000kb)

ER_ED_07_02_21_418000kb <- MERGED_07_02_21_418000kb %>% group_by(VAERS_ID, ER_ED_VISIT) %>% filter(ER_ED_VISIT == "Y")
nrow(ER_ED_07_02_21_418000kb)
```
```{r}
DEAD_Y_07_02_21_418000kb <- MERGED_07_02_21_418000kb %>% filter(DIED == "Y")
HOSPITAL_Y_07_02_21_418000kb <- MERGED_07_02_21_418000kb %>% filter(HOSPITAL == "Y")
ER_ED_Y_07_02_21_418000kb <- MERGED_07_02_21_418000kb %>% filter(ER_ED_VISIT == "Y")
L_THREAT_Y_07_02_21_418000kb <- MERGED_07_02_21_418000kb %>% filter(L_THREAT == "Y")
DISABLE_Y_07_02_21_418000kb <- MERGED_07_02_21_418000kb %>% filter(DISABLE == "Y")
BIRTH_DEFECT_Y_07_02_21_418000kb <- MERGED_07_02_21_418000kb %>% filter(BIRTH_DEFECT == TRUE)

nrow(DEAD_Y_07_02_21_418000kb)
nrow(HOSPITAL_Y_07_02_21_418000kb)
nrow(ER_ED_Y_07_02_21_418000kb)
nrow(L_THREAT_Y_07_02_21_418000kb)
nrow(DISABLE_Y_07_02_21_418000kb)
nrow(BIRTH_DEFECT_Y_07_02_21_418000kb)

SAES_07_02_21_418000kb <- rbind(DEAD_Y_07_02_21_418000kb, HOSPITAL_Y_07_02_21_418000kb, ER_ED_Y_07_02_21_418000kb, L_THREAT_Y_07_02_21_418000kb, DISABLE_Y_07_02_21_418000kb, BIRTH_DEFECT_Y_07_02_21_418000kb)

SAES_no_dups_07_02_21_418000kb <- subset(SAES_07_02_21_418000kb, !duplicated(subset(SAES_07_02_21_418000kb, select=c(VAERS_ID))))
nrow(SAES_no_dups_07_02_21_418000kb)
Percent = nrow(SAES_no_dups_07_02_21_418000kb)/nrow(MERGED_07_02_21_418000kb)*100
Percent
```
```{r}
SAES_no_dups_07_02_21_418000kb_AG <- SAES_no_dups_07_02_21_418000kb %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
SAES_no_dups_07_02_21_418000kb_AG_count <- SAES_no_dups_07_02_21_418000kb_AG %>% count(AGE_GROUP)
SAES_no_dups_07_02_21_418000kb_count_plot <- ggplot(SAES_no_dups_07_02_21_418000kb_AG_count, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("SEVERE ADVERSE EVENTS (N)") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("ALL SAEs -> Distribution of AGE GROUPS")
SAES_no_dups_07_02_21_418000kb_count_plot
ggsave("SAES_no_dups_07_02_21_418000kb_count_plot.png")

#MERGED_07_02_21_418000kb_AGE_count <- data.frame(MERGED_07_02_21_418000kb_AGE_count, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#MERGED_07_02_21_418000kb_AGE_count
```
BAR PLOT FOR SAE COMPARISON TO STANDARD
```{r}
Table_SAEs <- data.frame(
  Percent_SAE = c("Standard", "Actual"),
  SAE = as.numeric(c("58000", "66769")), #the first number is the SAE/total AE
  
  stringsAsFactors = FALSE
)
Table_SAEs_perc <- Table_SAEs %>% mutate("PERC" = round(SAE/(nrow(MERGED_07_02_21_418000kb))*100))
head(Table_SAEs_perc)

Table_SAEs_perc_plot <- ggplot(Table_SAEs_perc, aes(Percent_SAE, PERC, fill = Percent_SAE)) + geom_bar(stat = 'identity') + geom_text(aes(label = PERC)) + ylab("% VAERS SAE REPORTS") + theme(axis.text.y = element_text(colour = "#000000", size = 12)) + theme(axis.text.x = element_text(angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(legend.position = "none") + ggtitle("SAEs -> Comparison of Standard versus actual % reports")
Table_SAEs_perc_plot
```
BREAKTHROUGH COVID19 CASES
```{r}
strings_covid <- c("COVID-19", "Suspected COVID-19", "SARS-CoV-2 test positive", "COVID-19 pneumonia", "Asymptomatic COVID-19", "Vaccine breakthrough infection")
COVID_CASES <- function(SYMPTOM) {
  if(str_detect(SYMPTOM, paste(strings_covid, collapse = "|"))) {
      COVID19 = "COVID_CASE"
      } 
        else {
          COVID19 = "NA"
      }
    return(COVID19)
}
COVID_CASES1 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM1 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM1)) %>% filter(COVID_CASES != "NA")
COVID_CASES2 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM2 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM2)) %>% filter(COVID_CASES != "NA")
COVID_CASES3 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM3 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM3)) %>% filter(COVID_CASES != "NA")
COVID_CASES4 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM4 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM4)) %>% filter(COVID_CASES != "NA")
COVID_CASES5 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM5 != "NA") %>% mutate("COVID_CASES" = COVID_CASES(SYMPTOM5)) %>% filter(COVID_CASES != "NA")

COVID_CASES_Total <- rbind(COVID_CASES1,COVID_CASES2,COVID_CASES3,COVID_CASES4,COVID_CASES5)
COVID_CASES_Total <- data.frame(COVID_CASES_Total, stringsAsFactors = FALSE)

COVID_CASES_Total_ID <- COVID_CASES_Total %>% distinct(VAERS_ID, .keep_all = TRUE)
nrow(COVID_CASES_Total_ID)

COVID_CASES_Total_ENTRY <- COVID_CASES_Total_ID %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)
```
```{r}
#nrow(COVID_CASES_Total)
#nrow(COVID_CASES_Total_ID)
COVID_CASES <- COVID_CASES_Total_ID
#nrow(COVID_CASES)/nrow(MERGED_07_02_21_418000kb)*100
```
```{r}
AGE_GROUPS_COVID <- COVID_CASES %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
AGE_GROUPS_COVID_count <- AGE_GROUPS_COVID %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_COVID_plot <- ggplot(AGE_GROUPS_COVID_count, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF COVID-19 BREAKTHROUGH INFECTIONS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("Breakthough Covid-19 Infections -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_COVID_plot
ggsave("AGE_GROUPS_COVID_plot.png")

#AGE_GROUPS_COVID_plot_skew <- data.frame(AGE_GROUPS_COVID_count, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_COVID_plot_skew
```
```{r}
#nrow(COVID_CASES)
Table_IDs_COVID_percent <- data.frame(
  DATE = as.Date(c("1/16/21", "1/23/21", "1/30/21", "2/13/21", "2/27/21", "3/5/21", "3/12/21", "3/19/21", "3/26/21", "4/2/21", "4/9/21", "4/16/21", "4/23/21", "4/30/21", "5/7/21", "5/12/21", "5/19/21", "5/28/21", "6/4/21", "6/11/21", "6/18/21", "6/25/21"), "%m/%d/%y"),
  COVID = as.numeric(c("194", "254", "326", "619", "717", "776", "943", "1057", "1169", "1267", "1375", "1404", "1587", "1930", "2776", "3317", "3742", "4285", "4850", "5497", "6151", "9442")),
  
  stringsAsFactors = FALSE
)
Table_IDs_COVID_plot <- ggplot(data = Table_IDs_COVID_percent, aes(x = DATE, y = COVID)) + geom_line(data = Table_IDs_COVID_percent, aes(x = DATE, y = COVID, colour = "COVID"), colour = "red", size = 2) + geom_point(data = Table_IDs_COVID_percent, aes(DATE, COVID, colour = "COVID"), colour = "red", size = 5) + xlab("DATE") +  ylab("NUMBER OF BREAKTHROUGH INFECTIONS") + labs(title = "ABSOLUTE NUMBER OF BREAKTHROUGH INFECTIONS (CUMULATIVE)", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=15,vjust=0.5,hjust=0.5, size = 10)) + theme(panel.background = element_rect(fill = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + scale_x_continuous(breaks = round(seq(min(Table_IDs_COVID_percent$DATE), max(Table_IDs_COVID_percent$DATE), by = 10),0.1))
Table_IDs_COVID_plot
```
```{r}
COVID_CASES_count <- COVID_CASES %>% group_by(VAX_MANU, SYMPTOM1,SYMPTOM2,SYMPTOM3,SYMPTOM4,SYMPTOM5) %>% filter(VAX_MANU != "NA" & VAX_MANU != "UNKNOWN MANUFACTURER") %>% count(SYMPTOM1,SYMPTOM2,SYMPTOM3,SYMPTOM4,SYMPTOM5)
COVID_CASES_count_plot <- ggplot(COVID_CASES_count, aes(VAX_MANU, n, fill = VAX_MANU)) + geom_bar(stat = "Identity") + xlab("ID COUNT (N)") + ylab("COVID DIAGNOSIS") + theme(axis.text.y = element_text(colour = "#000000", size = 12)) + theme(axis.text.x = element_text(angle=0,vjust=0.5,hjust=0.5, size = 10)) + ggtitle("COVID CASES AS PER MANUFACTURER") + theme(legend.position = "none")
COVID_CASES_count_plot
```

```{r}
COVID_CASES_count_VM <- COVID_CASES %>% group_by(VAX_MANU) %>% count(VAX_MANU) %>% filter(VAX_MANU != "NA" & VAX_MANU != "UNKNOWN MANUFACTURER") %>% mutate("PERC" = n/(nrow(COVID_CASES))*100) %>% mutate("Round" = round(PERC, 2))
COVID_CASES_count_VM <- data.frame(COVID_CASES_count_VM, stringsAsFactors = FALSE) 
#COVID_CASES_count_VM
```
```{r}
strings_FRI <- c("Abnormal uterine bleeding", "Afterbirth pain", "Dysmenorrhoea", "Female reproductive tract disorder", "Female sex hormone level", "Female reproductive tract disorder", "Alpha 1 foetoprotein abnormal", "Heavy menstrual bleeding", "Intermenstrual bleeding", "Intra-uterine contraceptive device removal", "Menstrual disorder", "Menstruation delayed", "Menstruation irregular", "Premenstrual syndrome", "Abortion spontaneous", "Abortion spontaneous incomplete", "Uterine leiomyoma", "Uterine haemorrhage", "Uterine spasm", "Bradycardia foetal", "Biopsy foetal", "Bradycardia foetal", "Foetal cystic hygroma", "Foetal death", "Foetal exposure during pregnancy", "Foetal heart rate abnormal", "Foetal hypokinesia", "Foetal non-stress test abnormal", "Female sex hormone level", "Amenorrhoea", "Hypomenorrhoea", "Menopause", "Menorrhagia", "Oligomenorrhoea", "Polymenorrhoea", "Postmenopausal haemorrhage", "Postmenopause", "Vaginal discharge", "Vaginal disorder", "Vaginal erosion", "Vaginal haemorrhage", "Vaginal lesion", "Vaginal infection", "Vulvovaginal mycotic infection", "Vulval ulceration", "Vulval abscess", "Vulva cyst", "Cervix inflammation", "Adenocarcinoma of the cervix", "Cervix carcinoma", "Amniorrhoea", "Amniotic cavity infection", "Amniotic membrane rupture test positive", "Amniotic fluid index decreased")
```
```{r}
FRI <- function(SYMPTOM) {
  if(str_detect(SYMPTOM, paste(strings_FRI, collapse = "|"))) {
      FRI = "Female_Reprod_Issues"
      } 
        else {
          FRI = "NA"
      }
    return(FRI)
}
MERGED_07_02_21_418000kb_FRI1 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM1 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM1)) %>% filter(Female_Reprod_Issues != "NA")
MERGED_07_02_21_418000kb_FRI2 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM2 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM2)) %>% filter(Female_Reprod_Issues != "NA")
MERGED_07_02_21_418000kb_FRI3 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM3 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM3)) %>% filter(Female_Reprod_Issues != "NA")
MERGED_07_02_21_418000kb_FRI4 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM4 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM4)) %>% filter(Female_Reprod_Issues != "NA")
MERGED_07_02_21_418000kb_FRI5 <- MERGED_07_02_21_418000kb %>% rowwise() %>% filter(SYMPTOM5 != "NA") %>% mutate("Female_Reprod_Issues" = FRI(SYMPTOM5)) %>% filter(Female_Reprod_Issues != "NA")

MERGED_07_02_21_418000kb_FRI_Total <- rbind(MERGED_07_02_21_418000kb_FRI1,MERGED_07_02_21_418000kb_FRI2,MERGED_07_02_21_418000kb_FRI3,MERGED_07_02_21_418000kb_FRI4,MERGED_07_02_21_418000kb_FRI5)
MERGED_07_02_21_418000kb_FRI_Total <- data.frame(MERGED_07_02_21_418000kb_FRI_Total, stringsAsFactors = FALSE)

MERGED_07_02_21_418000kb_FRI_Total_ID <- MERGED_07_02_21_418000kb_FRI_Total %>% distinct(VAERS_ID, .keep_all = TRUE)
nrow(MERGED_07_02_21_418000kb_FRI_Total_ID)

MERGED_07_02_21_418000kb_FRI_Total_ENTRY <- MERGED_07_02_21_418000kb_FRI_Total_ID %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)

#nrow(MERGED_07_02_21_418000kb_FRI_Total_ENTRY)
MERGED_07_02_21_418000kb_FRI_Total_ENTRY <- data.frame(MERGED_07_02_21_418000kb_FRI_Total_ENTRY, stringsAsFactors = FALSE)
```
```{r}
#nrow(MERGED_07_02_21_418000kb_FRI1)
#nrow(MERGED_07_02_21_418000kb_FRI2)
#nrow(MERGED_07_02_21_418000kb_FRI3)
#nrow(MERGED_07_02_21_418000kb_FRI4)
#nrow(MERGED_07_02_21_418000kb_FRI5)
#MERGED_07_02_21_418000kb_FRI_Total
```
```{r}
#nrow(MERGED_07_02_21_418000kb_FRI_Total)
#nrow(MERGED_07_02_21_418000kb_FRI_Total_ID)
FRI_CASES <- MERGED_07_02_21_418000kb_FRI_Total
#nrow(FRI_CASES)/nrow(MERGED_07_02_21_418000kb)*100
```

```{r}
AGE_GROUPS_COVID_dead <- data.frame(AGE_GROUPS_COVID, stringsAsFactors = FALSE) %>% group_by(AGE_GROUP, DIED) %>% filter(DIED == "Y") %>% count(DIED) %>% mutate("SUM" = as.numeric(sum(n))) %>% mutate("Perc_D_tot" = as.numeric(n/(nrow(AGE_GROUPS_COVID))*100)) %>% mutate("Round" = round(Perc_D_tot, 2))
#AGE_GROUPS_COVID_dead
```
```{r}
no_COVID_CASES_count_VM <- COVID_CASES %>% group_by(VAX_MANU) %>% count(VAX_MANU) %>% filter(VAX_MANU != "NA" & VAX_MANU != "UNKNOWN MANUFACTURER") %>% mutate("PERC" = n/(nrow(MERGED_07_02_21_418000kb))*100) %>% mutate("Round" = round(PERC, 2))
#no_COVID_CASES_count_VM
COVID_CASES_count_VM <- COVID_CASES %>% group_by(VAX_MANU) %>% count(VAX_MANU) %>% filter(VAX_MANU != "NA" & VAX_MANU != "UNKNOWN MANUFACTURER") %>% mutate("PERC" = n/(nrow(COVID_CASES))*100) %>% mutate("Round" = round(PERC, 2))
COVID_CASES_count_VM <- data.frame(COVID_CASES_count_VM, stringsAsFactors = FALSE) 
#COVID_CASES_count_VM
```
```{r}
pie_VM <- plot_ly(COVID_CASES_count_VM, labels = ~VAX_MANU, values = ~Round, type = 'pie')
pie_VM <- pie_VM %>% layout(title = 'BREAKTHROUGH COVID-19 CASES BY MANUFACTURER',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
pie_VM
```
```{r}
pie_AGE_GROUPS_COVID_dead <- plot_ly(AGE_GROUPS_COVID_dead, labels = ~AGE_GROUP, values = ~Round, type = 'pie')
pie_AGE_GROUPS_COVID_dead <- pie_AGE_GROUPS_COVID_dead %>% layout(title = 'BREAKTHROUGH COVID-19 CASES RESULTING IN DEATH BY AGE GROUP',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
pie_AGE_GROUPS_COVID_dead
```

```{r}
FRI_CASES_ID <- FRI_CASES %>% group_by(AGE_YRS) %>% filter(AGE_YRS != "NA") %>% count(AGE_YRS)
FRI_CASES_ID_plot <- ggplot(FRI_CASES_ID, aes(AGE_YRS, n, fill = AGE_YRS)) + geom_bar(stat = "Identity") + xlab("AGE YEARS") + ylab("Number of Female Reproductive Issues") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14))  + theme(legend.position = "none") + ggtitle("Number of Female Reproductive Issues -> VAERS IDs vs AGE YEARS") + xlim(0,100)
FRI_CASES_ID_plot
ggsave("FRI_CASES_ID_plot.png")
```
```{r}
FRI_CASES <- data.frame(FRI_CASES, stringsAsFactors = FALSE) %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2020) 
FRI_CASES_VD_OD <- FRI_CASES %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) %>% arrange(desc(n)) %>% mutate("OBSERVED" = n) %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = as.integer(mean(n))) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(FRI_CASES)*100)
#head(FRI_CASES_VD_OD)

#chisq.test(FRI_CASES_VD_OD$OBSERVED)
```

```{r}
FRI_CASES_VD_OD_TILE_count <- data.frame(FRI_CASES, stringsAsFactors = FALSE) %>% group_by(VAX_DATE, ONSET_DATE) %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021) %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=25) %>% count(diff_in_days) %>% arrange(desc(n))  

FRI_CASES_VD_DD_TILE_plot <- ggplot(FRI_CASES_VD_OD_TILE_count, aes(VAX_DATE, ONSET_DATE)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("ONSET OF AEs") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.x = element_text(colour = "#000000", size = 14)) + theme(axis.text.y = element_text(colour = "#000000", size = 14), axis.text.x=element_text(angle=0,vjust=0.5,hjust=0.5)) + ggtitle("VACCINE DATE vs FRI AE ONSET DATE") 
FRI_CASES_VD_DD_TILE_plot 
ggsave("FRI_CASES_VD_DD_TILE_plot.png")
```

```{r}
CAUSATION_FRI_CASES_plot <- ggplot() + geom_line(data = FRI_CASES_VD_OD, aes(diff_in_days, Percentage_OBSERVED), colour = "orange", size = 2) + ylab("PERCENTAGE REPORTED FRI AEs") + xlab("NUMBER OF DAYS FOLLOWING INJECTION") + geom_line(data = FRI_CASES_VD_OD, aes(diff_in_days, Percentage_EXPECTED), colour = "yellow", size = 2) + labs(title = "FEMALE REPRODUCTIVE ISSUES post injection", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
CAUSATION_FRI_CASES_plot 
ggsave("CAUSATION_FRI_CASES_plot.png")
```

```{r}
D <- MERGED_07_02_21_418000kb %>% filter(DIED == "Y") %>% filter(year(DATEDIED) <= 2021 & year(DATEDIED) >= 2020) 
MERGED_07_02_21_418000kb_VD_DD <- D %>% mutate("diff_in_days" = difftime(DATEDIED, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) %>% arrange(desc(n)) %>% mutate("OBSERVED" = n) %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = as.integer(mean(n))) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(D)*100)
#head(MERGED_07_02_21_418000kb_VD_DD)

#chisq.test(MERGED_07_02_21_418000kb_VD_DD$OBSERVED)
```

```{r}
CAUSATION_DEATH_plot <- ggplot() + geom_line(data = MERGED_07_02_21_418000kb_VD_DD, aes(diff_in_days, Percentage_OBSERVED), colour = "orange", size = 2) + ylab("PERCENTAGE OF REPORTED DEATHS") + xlab("NUMBER OF DAYS FOLLOWING INJECTION") + geom_line(data = MERGED_07_02_21_418000kb_VD_DD, aes(diff_in_days, Percentage_EXPECTED), colour = "yellow", size = 2) + labs(title = "DEATHS post injection", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(panel.background = element_rect(fill = "#100000", color = "pink")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ylim(0,15)
CAUSATION_DEATH_plot 
ggsave("CAUSATION_DEATH_plot.png")
```

```{r}
VD_DD_TILE_count <- MERGED_07_02_21_418000kb %>% group_by(DIED, DATEDIED, VAX_DATE) %>% filter(DIED == "Y") %>% filter(year(DATEDIED) <= 2021 & year(DATEDIED) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021)%>% mutate("diff_in_days" = difftime(DATEDIED, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) 

#png("VD_DD_TILE_plot ", width = 4, height = 4, units = 'in', res = 100)

VD_DD_TILE_plot <- ggplot(VD_DD_TILE_count, aes(VAX_DATE, DATEDIED)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("DATE OF DEATH") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + ggtitle("VACCINE DATE vs DEATH DATE OF AEs") 
VD_DD_TILE_plot 
ggsave("VD_DD_TILE_plot.png")
```
```{r}
H <- MERGED_07_02_21_418000kb %>% filter(HOSPITAL == "Y") %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2020)
MERGED_07_02_21_418000kb_H_VD_OD <- H %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) %>% arrange(desc(n))  %>% mutate("OBSERVED" = n)  %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = mean(n)) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(H)*100)
#head(MERGED_07_02_21_418000kb_H_VD_OD)

#chisq.test(MERGED_07_02_21_418000kb_H_VD_OD$OBSERVED)
```

```{r}
CAUSATION_HOSP_plot <- ggplot() + geom_line(data = MERGED_07_02_21_418000kb_H_VD_OD, aes(diff_in_days, Percentage_OBSERVED), colour = "orange", size = 2) + xlab("NUMBER OF DAYS FOLLOWING INJECTION") +  ylab("PERCENTAGE OF REPORTED HOSPITALIZATIONS") + geom_line(data = MERGED_07_02_21_418000kb_H_VD_OD, aes(diff_in_days, Percentage_EXPECTED), colour = "yellow", size = 2) + labs(title = "HOSPITALIZATIONS post injection", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000", color = "pink")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
CAUSATION_HOSP_plot
ggsave("CAUSATION_HOSP_plot.png")
```

```{r}
H_VD_OD_TILE_count <- MERGED_07_02_21_418000kb %>% group_by(HOSPITAL, ONSET_DATE, VAX_DATE) %>% filter(HOSPITAL == "Y") %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021)%>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) 

H_VD_DD_TILE_plot <- ggplot(H_VD_OD_TILE_count, aes(VAX_DATE, ONSET_DATE)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("ONSET OF AEs") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.x = element_text(colour = "#000000", size = 14)) + theme(axis.text.y = element_text(colour = "#000000", size = 14), axis.text.x=element_text(angle=0,vjust=0.5,hjust=0.5)) + ggtitle("VACCINE DATE vs HOSPITALIZATION DATE") 
H_VD_DD_TILE_plot 
ggsave("H_VD_DD_TILE_plot.png")
```
```{r}
ER <- MERGED_07_02_21_418000kb %>% filter(ER_ED_VISIT == "Y") %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2020)
MERGED_07_02_21_418000kb_ER_VD_OD <- ER %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) %>% arrange(desc(n))  %>% mutate("OBSERVED" = n) %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = mean(n)) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(ER)*100)
#head(MERGED_07_02_21_418000kb_ER_VD_OD)

#chisq.test(MERGED_07_02_21_418000kb_ER_VD_OD$OBSERVED)
```

```{r}
CAUSATION_ER_plot <- ggplot() + geom_line(data = MERGED_07_02_21_418000kb_ER_VD_OD, aes(diff_in_days, Percentage_OBSERVED), colour = "orange", size = 2) + xlab("NUMBER OF DAYS FOLLOWING INJECTION") +  ylab("PERCENTAGE OF REPORTED ER VISITS") + geom_line(data = MERGED_07_02_21_418000kb_ER_VD_OD, aes(diff_in_days, Percentage_EXPECTED), colour = "yellow", size = 2) + labs(title = "ER VISITS post injection", caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000", color = "pink")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
CAUSATION_ER_plot 
ggsave("CAUSATION_ER_plot.png")
```

```{r}
ER_VD_OD_TILE_count <- MERGED_07_02_21_418000kb %>% group_by(ER_ED_VISIT, ONSET_DATE, VAX_DATE) %>% filter(ER_ED_VISIT == "Y") %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021)%>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) 

ER_VD_DD_TILE_plot <- ggplot(ER_VD_OD_TILE_count, aes(VAX_DATE, ONSET_DATE)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("ONSET OF AEs") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.x = element_text(colour = "#000000", size = 14)) + theme(axis.text.y = element_text(colour = "#000000", size = 14), axis.text.x=element_text(angle=0,vjust=0.5,hjust=0.5)) + ggtitle("VACCINE DATE vs ER visit DATE") 
ER_VD_DD_TILE_plot 
ggsave("ER_VD_DD_TILE_plot.png")
```

```{r}
AGE_GROUPS_D <- D %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
AGE_GROUPS_H <- H %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
AGE_GROUPS_ER <- ER %>% rowwise() %>% filter(AGE_YRS != "NA") %>% mutate("AGE_GROUP" = AGE_GROUPED_BY_DECADE(AGE_YRS))
```

```{r}
AGE_GROUPS_D_ALL_ID <- AGE_GROUPS_D %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_D_ALL_ID_plot <- ggplot(AGE_GROUPS_D_ALL_ID, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF DEATHS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("Death -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_D_ALL_ID_plot
ggsave("AGE_GROUPS_D_ALL_ID_plot.png")
#AGE_GROUPS_D_ALL_ID_skew <- data.frame(AGE_GROUPS_D_ALL_ID, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_D_ALL_ID_skew
```
```{r}
AGE_GROUPS_H_ALL_ID <- AGE_GROUPS_H %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_H_ALL_ID_plot <- ggplot(AGE_GROUPS_H_ALL_ID, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF HOSPITALIZATIONS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("Hospital -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_H_ALL_ID_plot
ggsave("AGE_GROUPS_H_ALL_ID_plot.png")
#AGE_GROUPS_H_ALL_ID_skew <- data.frame(AGE_GROUPS_H_ALL_ID, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_H_ALL_ID_skew
```
```{r}
AGE_GROUPS_ER_ALL_ID <- AGE_GROUPS_ER %>% group_by(AGE_GROUP) %>% filter(AGE_GROUP != "NA") %>% count(AGE_GROUP)
AGE_GROUPS_ER_ALL_ID_plot <- ggplot(AGE_GROUPS_ER_ALL_ID, aes(AGE_GROUP, n, fill = AGE_GROUP)) + geom_bar(stat = "Identity") + geom_text(aes(label = n)) + xlab("AGE GROUPS") + ylab("NUMBER OF EMERGENCY DOCTOR VISITS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=25,vjust=0.5,hjust=0.5, size = 8)) + theme(legend.position = "none") + ggtitle("ER -> VAERS IDs vs AGE GROUPS")
AGE_GROUPS_ER_ALL_ID_plot
ggsave("AGE_GROUPS_ER_ALL_ID_plot.png")
#AGE_GROUPS_ER_ALL_ID_skew <- data.frame(AGE_GROUPS_ER_ALL_ID, stringsAsFactors = FALSE) %>% mutate("skew" = 3*(mean(n)-median(n))/(sd(n)))
#AGE_GROUPS_ER_ALL_ID_skew
```

```{r}
Table_IDs_percent <- data.frame(
  DATE = as.Date(c("1/16/21", "1/23/21", "1/30/21", "2/13/21", "2/27/21", "3/5/21", "3/12/21", "3/19/21", "3/26/21", "4/2/21", "4/9/21", "4/16/21", "4/23/21", "4/30/21", "5/7/21", "5/14/21", "5/21/21", "5/28/21", "6/4/21", "6/11/21", "6/18/21", "6/25/21", "7/2/21"), "%m/%d/%y"),
  IDs = as.numeric(c("1431", "2160", "2946", "5351", "9286", "14701", "20586", "27955", "34121", "40348", "46163", "57641", "75370", "108191", "146622", "182559", "217716", "252523", "284041", "316929", "346967", "368162", "388900")),
  FV = as.numeric(c("4582089", "7664179", "11037313", "18895522", "27167910", "31720149", "35000000", "44145522", "48700000", "57980000", "64420000", "80610000", "89250000", "99670000", "108930000", "118990000", "126610000", "132770000", "136640000", "141580000", "147760000", "147760000", "155880000")),
  US_population = as.numeric(c("330175152" , "330175733" , "330176314", "330176895", "330177476", "330178057", "330178638", "330179219", "330179800", "330180381", "330193928", "330209443", "330244315", "330239326", "330757983", "332700813", "332738974", "332371755", "332393075", "332414436", "332888350", "332888350", "332962963")),
  DEATH = as.numeric(c("137", "281", "456", "810", "984", "1162", "1419", "1561", "1957", "2149", "2240", "2500", "3084", "3442", "3731", "4015", "4169", "4372", "4547", "4700", "4825", "4931", "5163")),
  HOSPITAL = as.numeric(c("338", "607", "953", "1747", "2195", "2676", "3412", "3913", "4387", "4758", "4906", "5009", "6215", "8099", "10648", "11557", "12625", "14977", "17520", "19354", "20571", "21636", "22612")), 
  ER = as.numeric(c("338", "576", "847", "1493", "2187", "2980", "3855", "4763", "5529", "6329", "6983", "8084", "10375", "14553", "19620", "23641", "27774", "32547", "37184", "41937", "45952", "49511", "52399")),
  SAE = as.numeric(c("690", "1207", "1806", "3221", "4291", "5535", "7065", "8470", "9796", "10930", "11724", "13057", "16478", "22145", "29024", "33791", "38650", "44860", "51068", "57073", "61917", "66255", "66769")),
  COVID = as.numeric(c("194", "254", "326", "619", "717", "776", "943", "1057", "1169", "1267", "1375", "1404", "1587", "1930", "2776", "3317", "3742", "4285", "4850", "5497", "6151", "6414", "9442")), 
  FRI = as.numeric(c("4", "7", "8", "24", "34", "43", "67", "77", "88", "98", "105", "123", "165", "287", "436", "597", "774", "1067", "1421", "1967", "2460", "2940", "5351")),
  
  stringsAsFactors = FALSE
)
#Table_IDs_percent
```

```{r}
SAE_calculations <- Table_IDs_percent %>% mutate("SAE_p_FV" = SAE/(FV/1000000)) %>% mutate("SAE_p_mill_doses" = SAE_p_FV/(2.1)) #cum num of doses/FV population = 328150000/155880000 = 2.1
#SAE_calculations
```

```{r}
IDs = as.numeric(c("1431", "2160", "2946", "5351", "9286", "14701", "20586", "27955", "34121", "40348", "46163", "57641", "75370", "108191", "146622", "182559", "217716", "252523", "284041", "316929", "346967", "368162", "388900"))
#(diff(IDs)/(IDs))*100
```

```{r}
Table_IDs_percent_FV_IDs <- Table_IDs_percent %>% mutate("DEATH_PERCENT_ID" = DEATH/IDs*100) %>% mutate("H_PERCENT_ID" = HOSPITAL/IDs*100) %>% mutate("ER_PERCENT_ID" = ER/IDs*100) %>% mutate("SAE_PERCENT_ID" = SAE/IDs*100) %>% mutate("COVID_PERCENT_ID" = COVID/IDs*100) %>%  mutate("DEATH_PERCENT_FV" = DEATH/FV*100) %>% mutate("H_PERCENT_FV" = HOSPITAL/FV*100) %>% mutate("ER_PERCENT_FV" = ER/FV*100) %>% mutate("SAE_PERCENT_FV" = SAE/FV*100) %>% mutate("COVID_PERCENT_FV" = COVID/FV*100) %>% mutate("FRI_PERCENT_FV" = FRI/FV*100)
#Table_IDs_percent_FV_IDs
```

```{r} 
IDs_normalized_total_IDs <- Table_IDs_percent %>% mutate("IDs_normal_ID" = IDs/nrow(MERGED_07_02_21_418000kb)*100)
#IDs_normalized_total_IDs
```

```{r}
Table_IDs_percent_normalized_US_pop <- Table_IDs_percent_FV_IDs %>% mutate("IDs_norm_US_pop" = IDs/(US_population/1000000)) 
#Table_IDs_percent_normalized_US_pop

Table_IDs_percent_normalized_FV <- Table_IDs_percent_FV_IDs %>% mutate("IDs_norm_FV" = IDs/(FV/1000000)) 
#Table_IDs_percent_normalized_FV

Table_AEs_percent_normalized_per_million <- Table_IDs_percent_FV_IDs %>% mutate("Death_norm_per_million" = DEATH/(FV/1000000)) %>% mutate("Hospital_norm_per_million" = HOSPITAL/(FV/1000000)) %>% mutate("ER_norm_per_million" = ER/(FV/1000000)) %>% mutate("SAE_norm_per_million" = SAE/(FV/1000000)) %>% mutate("COVID_norm_per_million" = COVID/(FV/1000000)) %>% mutate("FRI_norm_per_million" = FRI/(FV/1000000))
#Table_AEs_percent_normalized_per_million
```

```{r}
Table_IDs_all_VAERS <- data.frame(
  DATE = as.double(c("1990", "1991", "1992", "1993", "1994", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021")),
  IDs = as.numeric(c("2787", "16657", "18392", "17655", "18441", "17794", "17592", "17247", "15997", "20072", "23113", "25492", "26852", "31248", "28663", "28664", "31353", "48349", "51019", "50524", "49601", "41200", "41149", "43335", "47165", "57224", "60450", "52751", "62305", "61204", "55833", "62000")),
  US_Pop = as.numeric(c("252120309", "254539370", "256990613", "259532129", "262241196", "265163745", "268335003" ,"271713635", "275175301", "278548150", "281710909", "284607993", "287279318", "289815562", "292354658", "294993511", "297758969", "300608429", "303486012", "306307567", "309011475", "311584047", "314043885", "316400538", "318673411", "320878310" ,"323015995", "325084756", "327096265", "329064917", "331002651", "332915073")),
  
  stringsAsFactors = FALSE
)
Table_IDs_all_VAERS <- Table_IDs_all_VAERS %>% mutate("NORMALIZED" = IDs/(US_Pop/1000000))
#head(Table_IDs_all_VAERS)
```
```{r}
Table_IDs_all_VAERS_plot <- ggplot(data = Table_IDs_all_VAERS, aes(DATE, IDs, colour = "IDs")) + geom_line(data = Table_IDs_all_VAERS, aes(DATE, IDs, colour = "IDs"), colour = "red", size = 2) + geom_point(data = Table_IDs_all_VAERS, aes(DATE, IDs, colour = "IDs"), colour = "red", size = 5) + xlab("DATE") + ylab("ABSOLUTE NUMBER OF VAERS IDs (N)") + labs(title = "Change in absolute number of VAERS IDs per year", caption = "Data source: VAERS", size = 12) + theme(axis.text.y = element_text(colour = "#000000", size = 14)) + theme(axis.text.x = element_text(angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + theme(legend.background = element_rect(fill = "white"), legend.key = element_rect(fill = "black", color = NA), legend.key.size = unit(0.5, "cm"), legend.key.width = unit(0.5,"cm")) + geom_smooth(method = "lm", se=FALSE, color="blue", aes(group=1))
#Table_IDs_all_VAERS_plot
#ggsave("Table_IDs_all_VAERS_plot.png")
```
```{r}
Table_IDs_all_VAERS_plot_norm_US <- ggplot(data = Table_IDs_all_VAERS, aes(DATE, NORMALIZED, colour = "NORMALIZED")) + geom_line(data = Table_IDs_all_VAERS, aes(DATE, NORMALIZED, colour = "NORMALIZED"), colour = "red", size = 2) + geom_point(data = Table_IDs_all_VAERS, aes(DATE, NORMALIZED, colour = "NORMALIZED"), colour = "red", size = 5) + xlab("DATE") +  ylab("VAERS ID rate normalized to U.S. popln/million") + labs(title = "VAERS ID counts for the past 30 years", caption = "Data source: VAERS", size = 12) + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + theme(legend.background = element_rect(fill = "white"), legend.key = element_rect(fill = "black", color = NA), legend.key.size = unit(0.5, "cm"), legend.key.width = unit(0.5,"cm")) + geom_smooth(method = "lm", se=FALSE, color="blue", aes(group=1)) #+ geom_text(aes(label = IDs), color = "white", size = 4) 
Table_IDs_all_VAERS_plot_norm_US
ggsave("Table_IDs_all_VAERS_plot_norm_US.png")
```
```{r}
Table_IDs_plot <- ggplot(data = Table_IDs_percent, aes(x = DATE, y = IDs)) + geom_line(data = Table_IDs_percent, aes(x = DATE, y = IDs, colour = "IDs"), colour = "red", size = 2) + geom_point(data = Table_IDs_percent, aes(DATE, IDs, colour = "IDs"), colour = "red", size = 5) + xlab("DATE") +  ylab("Absolute number of VAERS IDs (N)") + labs(caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=15,vjust=0.5,hjust=0.5, size = 10)) + theme(panel.background = element_rect(fill = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + scale_x_continuous(breaks = round(seq(min(Table_IDs_percent$DATE), max(Table_IDs_percent$DATE), by = 10),0.1)) #+ ylim(0,200000)
Table_IDs_plot
ggsave("Table_IDs_plot.png")
```

```{r}
Table_IDs_plot_FV_p_mill <- ggplot(data = Table_IDs_percent_normalized_FV, aes(x = DATE, y = IDs_norm_FV)) + geom_line(data = Table_IDs_percent_normalized_FV, aes(x = DATE, y = IDs_norm_FV, colour = "IDs_norm_FV"), colour = "red", size = 2) + geom_point(data = Table_IDs_percent_normalized_FV, aes(DATE, IDs_norm_FV, colour = "IDs_norm_FV"), colour = "red", size = 5) + xlab("DATE") +  ylab("VAERS IDs per million FV") + labs(caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=15,vjust=0.5,hjust=0.5, size = 10)) + theme(panel.background = element_rect(fill = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + scale_x_continuous(breaks = round(seq(min(Table_IDs_percent$DATE), max(Table_IDs_percent$DATE), by = 10),0.1)) #+ ylim(0,200000)
Table_IDs_plot_FV_p_mill
ggsave("Table_IDs_plot_FV_p_mill.png")
```
```{r}
Table_IDs_percent_plot <- ggplot() + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, DEATH, colour = "DEATH"), size = 2) + xlab("DATE") +  ylab("Absolute number of VAERS IDs (N)") + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, HOSPITAL, colour = "HOSPITAL"), size = 2) + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, ER, colour = "ER"), size = 2) + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, SAE, colour = "SAE"), size = 2) + xlab("DATE") +  ylab("Absolute number of VAERS IDs (N)") + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, COVID, colour = "COVID"), size = 2) + geom_line(data = Table_IDs_percent_FV_IDs, aes(DATE, FRI, colour = "FRI"), size = 2) + labs(caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + theme(legend.background = element_rect(fill = "white"), legend.key = element_rect(fill = "black", color = NA), legend.key.size = unit(0.5, "cm"), legend.key.width = unit(0.5,"cm")) 
Table_IDs_percent_plot
ggsave("Table_IDs_percent_plot.png")
```

```{r}
Table_IDs_FV_per_million <- ggplot() + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, Death_norm_per_million, colour = "DEATH"), size = 2) + xlab("DATE") +  ylab("AEs/million") + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, Hospital_norm_per_million, colour = "HOSPITAL"), size = 2) + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, ER_norm_per_million, colour = "ER"), size = 2) + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, SAE_norm_per_million, colour = "SAE"), size = 2)+ geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, COVID_norm_per_million, colour = "COVID"), size = 2) + geom_line(data = Table_AEs_percent_normalized_per_million, aes(DATE, FRI_norm_per_million, colour = "FRI"), size = 2) + labs(caption = "Data source: VAERS") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(panel.background = element_rect(fill = "#100000")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.title=element_blank()) + theme(legend.background = element_rect(fill = "white"), legend.key = element_rect(fill = "black", color = NA), legend.key.size = unit(0.5, "cm"), legend.key.width = unit(0.5,"cm"))
Table_IDs_FV_per_million
ggsave("Table_IDs_FV_per_million.png")
```
```{r}
MERGED_07_02_21_418000kb_12_18 <- MERGED_07_02_21_418000kb %>% filter(AGE_YRS >=12 & AGE_YRS <=18)
nrow(MERGED_07_02_21_418000kb_12_18)

MERGED_07_02_21_418000kb_12_18_1 <- MERGED_07_02_21_418000kb_12_18 %>% distinct(VAERS_ID, .keep_all = TRUE)
YOUTH_CASES <- MERGED_07_02_21_418000kb_12_18_1 %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)

YOUTH_CASES_ID <- YOUTH_CASES %>% group_by(AGE_YRS) %>% filter(AGE_YRS != "NA") %>% count(AGE_YRS)
YOUTH_CASES_ID_plot <- ggplot(YOUTH_CASES_ID, aes(AGE_YRS, n, fill = AGE_YRS)) + geom_bar(stat = "Identity") + xlab("AGE YEARS") + ylab("Number of YOUTH 12-15 y/o CASES") + theme(axis.text.y = element_text(colour = "black", size = 14)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 14))  + theme(legend.position = "none") + ggtitle("Number of AEs in 12-15 y/os  -> VAERS IDs vs AGE YEARS") #+ xlim(0,100)
YOUTH_CASES_ID_plot
ggsave("YOUTH_CASES_ID_plot.png")

YOUTH_CASES <- data.frame(YOUTH_CASES, stringsAsFactors = FALSE) %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2020) 
YOUTH_CASES_VD_OD <- YOUTH_CASES %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=50) %>% count(diff_in_days) %>% arrange(desc(n)) %>% mutate("OBSERVED" = n) %>% mutate("Percentage_OBSERVED" = n/sum(n)*100) %>% mutate("EXPECTED" = as.integer(mean(n))) %>% mutate("Percentage_EXPECTED" = EXPECTED/nrow(FRI_CASES)*100)
#head(YOUTH_CASES_VD_OD)

#chisq.test(YOUTH_CASES_VD_OD$OBSERVED)
```

```{r}
YOUTH_CASES_VD_OD_TILE_count <- data.frame(YOUTH_CASES, stringsAsFactors = FALSE) %>% group_by(VAX_DATE, ONSET_DATE) %>% filter(year(ONSET_DATE) <= 2021 & year(ONSET_DATE) >= 2021) %>% filter(year(VAX_DATE) <= 2021 & year(VAX_DATE) >= 2021) %>% mutate("diff_in_days" = difftime(ONSET_DATE, VAX_DATE, units = "days")) %>% filter(diff_in_days >= 0 & diff_in_days <=25) %>% count(diff_in_days) %>% arrange(desc(n))  

YOUTH_CASES_VD_DD_TILE_plot <- ggplot(YOUTH_CASES_VD_OD_TILE_count, aes(VAX_DATE, ONSET_DATE)) + geom_tile(aes(fill = n)) + scale_fill_gradient('n', low = "blue", high = "red") + theme(panel.background = element_rect(fill = "black", color = 'black')) + ylab("ONSET OF AEs") + xlab("VACCINE DATE") + theme(panel.grid=element_blank()) + theme(axis.text.x = element_text(colour = "#000000", size = 14)) + theme(axis.text.y = element_text(colour = "#000000", size = 14), axis.text.x=element_text(angle=0,vjust=0.5,hjust=0.5)) + ggtitle("VACCINE DATE vs YOUTH CASES 12-15 y/o AE ONSET DATE") 
YOUTH_CASES_VD_DD_TILE_plot 
ggsave("FYOUTH_CASES_VD_DD_TILE_plot.png")
```
```{r}
MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo1 <- MERGED_07_02_21_418000kb %>% group_by(VAERS_ID, SYMPTOM1) %>% filter(SYMPTOM1 == "Myocarditis") %>% arrange(desc(AGE_YRS))
#nrow(MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo1)
MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo2 <- MERGED_07_02_21_418000kb %>% group_by(VAERS_ID, SYMPTOM2) %>% filter(SYMPTOM2 == "Myocarditis") %>% arrange(desc(AGE_YRS))
#nrow(MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo2)
MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo3 <- MERGED_07_02_21_418000kb %>% group_by(VAERS_ID, SYMPTOM3) %>% filter(SYMPTOM3 == "Myocarditis") %>% arrange(desc(AGE_YRS))
#nrow(MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo3)
MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo4 <- MERGED_07_02_21_418000kb %>% group_by(VAERS_ID, SYMPTOM4) %>% filter(SYMPTOM4 == "Myocarditis") %>% arrange(desc(AGE_YRS))
#nrow(MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo4)
MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo5 <- MERGED_07_02_21_418000kb %>% group_by(VAERS_ID, SYMPTOM5) %>% filter(SYMPTOM5 == "Myocarditis") %>% arrange(desc(AGE_YRS))
#nrow(MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo5)
myo_Total <- rbind(MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo1,MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo2,MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo3,MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo4,MERGED_SYM_DAT_VAX_07_02_21_418000kb_myo5)
myo_Total <- data.frame(myo_Total, stringsAsFactors = FALSE)
nrow(myo_Total)
nrow(myo_Total)/nrow(MERGED_07_02_21_418000kb)*100
#myo_Total
myo_Total_count <- myo_Total %>% count(SYMPTOM1,SYMPTOM2,SYMPTOM3,SYMPTOM4,SYMPTOM5) %>% mutate("SUM" = sum(n)) %>% arrange(desc(n))
#myo_Total_count

myo_Total_ID <- myo_Total %>% distinct(VAERS_ID, .keep_all = TRUE)

myo_Total_ENTRY <- myo_Total_ID %>% distinct(AGE_YRS, STATE, SEX, VAX_LOT, VAX_SITE, VAX_ROUTE, VAX_DATE, VAX_NAME, VAX_MANU, RECVDATE, CAGE_YR, CAGE_MO, DIED, HOSPITAL, ER_ED_VISIT, ALLERGIES, ONSET_DATE, PRIOR_VAX, VAX_DOSE_SERIES, L_THREAT, DISABLE,	BIRTH_DEFECT, .keep_all = TRUE)

#nrow(myo_Total)
#nrow(myo_Total_ENTRY)

myo_Total_doses_all <- myo_Total  %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n))
#myo_Total_doses_all

myo_Total_doses_1_2 <- myo_Total  %>% filter(VAX_DOSE_SERIES <= 2) %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n))
#myo_Total_doses_1_2

myo_Total_doses_1_2_age <- myo_Total  %>% filter(VAX_DOSE_SERIES <= 2) %>% group_by(AGE_YRS) %>% count(VAX_DOSE_SERIES) %>% arrange(desc(n))
#myo_Total_doses_1_2_age 
```
```{r}
AGE_YEARS_myo_Total_plot <- ggplot(myo_Total_doses_1_2_age, aes(AGE_YRS, n, fill = VAX_DOSE_SERIES)) + geom_bar(stat = "identity", position = position_dodge(preserve = 'single')) + xlab("Age of reported patient, years") + ylab("Absolute number of Myocarditis") + labs(caption = "Data source: VAERS/Analysis: Dr. Jessica Rose") + theme(axis.text.y = element_text(colour = "black", size = 12)) + theme(axis.text.x = element_text(colour = "black", angle=0,vjust=0.5,hjust=0.5, size = 12)) + theme(legend.position = "right") + ggtitle("Myocarditis in VAERS after mRNA injection by age and dose #") + theme(legend.text=element_text(size=6), legend.title=element_text(size=6)) + xlim(12,62) #+ ylim(0,20)  
AGE_YEARS_myo_Total_plot
```
```
